---
title: "Credit Allocation Mix in Peer-to-Peer Lending: A Network Study on Bondora"
subtitle: "Social Network Analysis - Group 7"
date: 2025-11-16
author:
  - name: Floris 	Vermeulen
    affiliations:
      - ref: jads
      - ref: tiu
  - name: Martijn van Iterson
    affiliations:
      - ref: jads
      - ref: tiu
  - name: Niek	Fleerakkers
    affiliations:
      - ref: jads
      - ref: tiu
  - name: Patryk	Grodek
    affiliations:
      - ref: jads
      - ref: tue
  - name: Samir Sabitli
    affiliations:
      - ref: jads
      - ref: tue
affiliations:
  - id: jads
    name: Jheronimus Academy of Data Science
  - id: tiu
    name: Tilburg University
  - id: tue
    name: Eindhoven University of Technology
include-in-header:
  - text: |
      \usepackage{tgschola}
      \usepackage[T1]{fontenc}
      
      \usepackage{longtable}
      \usepackage{graphicx}
      \usepackage{booktabs}
      \usepackage{dcolumn}
      \usepackage{float}
      \usepackage{listings}
      \usepackage{color}
      \linespread{1.15}
      \usepackage{sectsty}
      \allsectionsfont{\sffamily}
bibliography: resources/r-references.bib
csl: resources/apa-6th-edition.csl
toc: true
link-citations: true
toc-title: "Table of Contents"
toc-depth: 2
format:
  pdf: 
    pdf-engine: xelatex
    fig-pos: H
    lof: true
    lot: true
    documentclass: article
    pagesize: a4
    number-sections: true
    df-print: kable
    fig-cap-location: top
    mainfont: "TeX Gyre Schola"
    mathfont: "Latin Modern Math"
    sansfont: "Arial"
    geometry:
      - left=1.15in
      - right=1.15in
      - bottom=1.15in
      - top=1.15in
      - heightrounded
execute:
    echo: false
    cache: false
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| label: analysis-preferences
#| include: false

# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
#install.packages("here")
```

```{r}
# Include any relevant objects
net <- readRDS(here::here("resources","objects","preprocessing",
                          "bondora_net.Rds"))
net_plot <- readRDS(here::here("resources","objects","ergm",
                          "network_plot.Rds"))
sum_tab <- readRDS(here::here("resources","objects","preprocessing",
                          "summary_table.Rds"))
cent_plots <- readRDS(here::here("resources","objects","ergm",
                          "network_plots.Rds"))
```

{{< pagebreak >}}

## Introduction {#sec-intro}

With the advent of online peer-to-peer (P2P) lending platforms, the
traditional methods of financial intermediation have been usurped by
individual choice; with no stringent third parties involved in
transactions, individuals have greater power in sourcing credit. This
presents new realities for individuals disenfranchised by the
traditional financial system [@alistair_milne_business_2017].
Traditional financial literature shows that individually, investors
exhibit behavioural biases [@agrawal_managing_2012; @ayal_two_2011] in
how they choose their investments, however, numerous studies demonstrate
that these behavioural biases are present among non-professional P2P
borrowers. For example, @herzenstein_strategic_2011 and
@lee_herding_2012 suggest that users cluster around popular loans,
exhibiting 'herding' behaviour. Literature focuses predominantly on
funding success [@yao_relationship_2019], however, @ayal_behavioral_2018
suggests that there is insufficient attention on how borrowers form
their loan portfolios. As the authors indicate, this area of research is
important in understanding why less professional investors deviate from
known models of rational investor behaviour. Our study aims to model and
understand how P2P borrowers' behaviours influence what kinds of loans
they acquire. In subsequent discussions, we consider investors as
rational participants in credit markets, as described by theory, whereas
borrowers refer to the users of P2P platforms.

Investigating behavioural biases, @ayal_behavioral_2018 find that
investors' familiarity with specific assets can make them seem less
risky, making recognisable assets more attractive. However, such assets
lead to undiversified portfolios. Evidence with P2P platforms supports
the familiarity bias in lenders, as @galak_micro-finance_2010
demonstrate that lenders choose borrowers who are similar across
demographic attributes such as gender, occupation, and ethnicity.
Moreover, @amar_winning_2011 explains that borrowers show an aversion to
holding a *large number of debts,* preferring to first reduce debt
count, even if they had other debts with higher interest rates. This
irrational behaviour demonstrates that individual borrowers face
behavioural biases on the *type* and *number* of debts they select.
Based on these notions, we formulate our first study, asking:

***Research Question 1:** Do borrowers vary in their loan use type and
frequency based on their individual and loan characteristics?*

This question is answered using linear Quadratic Assignment Procedure
(QAP) regressions, as we can represent similarities across loan uses and
attributes as matrices, weighted by loan use frequency, as inputs in the
model.

Using data from LendingClub, @serrano-cinca_determinants_2015 study
funding success, finding that the reported loan use is a significant
factor explaining differences in default rates. Specifically, the
authors argue that there is heterogeneity in the credit riskiness of
individuals seeking specific loan purposes, indicating that credit
rating can explain some variation in borrowers' choice of loan use.
Therefore, we formulate the following hypothesis:

*Hypothesis 1: Borrowers with the same credit ratings choose similar
loan uses*

Supporting the familiarity bias, @ravina_love_2019 studies personal
characteristics in P2P markets, finding that loans tend to be awarded to
those sharing personal characteristics such as occupation. As
@ayal_behavioral_2018 implies, this can be generalised to suggest that
lenders' loan grants reflect a wider network of similar borrower
preferences, which can reflect the homogeneity in loan use observed by
@serrano-cinca_determinants_2015. Thus,

*Hypothesis 2: Borrowers sharing the same occupation area tend to share
loan uses*

Further, behavioural finance suggests that individuals' financial
decisions are not independent, but rather, shaped by bounded rationality
and psychological preferences [@kahneman_prospect_1979]. In the P2P
context, @ayal_behavioral_2018 shows that these preferences can
influence how borrowers choose loan types and how this can overlap with
other borrowers' decisions. Using the Exponential Random Graph Model
(ERGM) framework, we can identify how these behaviours manifest as
structural regularities in a borrower-to-loan-use network. Therefore, we
ask:

***Research Question 2:** What behavioural mechanisms affect the
structure of borrower-to-loan-use connections?*

The aforedescribed section outlined that the familiarity bias can lead
to similarities across borrower portfolios in terms of their
characteristics. However, in P2P settings, this can manifest itself as
familiarity *clustering*, wherein borrowers who have previously been
granted certain loan types are more likely to repeat their choices, or
select other categories that appear familiar to different borrowers.
@herzenstein_strategic_2011, studying borrower decisions on Prosper.com,
identifies a 'strategic herding' behaviour, where individuals gravitate
towards popular or socially-validated loans. Considering that loan
popularity is directly observable through its number of bids, we assert
that borrowers linked to one loan purpose are likely to connect with
others sharing the same loan use. In network terms, this behaviour
results in cyclical substructures where there are overlapping nodes of
each of the bipartite partitions. We capture a conservative form of this
cycle through the ERGM term `cycle(4)`, formulating:

*Hypothesis 3: Borrowers who share one loan purpose are likely to share
another*

Further, @ayal_perceived_2009, expands on the theory of Debt Account
Aversion with the Theory of Perceived Diversification, explaining that
borrowers' distress increases with not just by the number of debts, but
also the perceived *distinctiveness* of each debt type, suggesting that
in the P2P context, debt mentally differentiated by purpose would also
be seen as distressing. Therefore, we assert that borrowers try to
minimise loan distinctiveness by requesting loans of a single type. This
can be captured by the `b1degree(1)` term, which models whether
borrowers are only linked to one loan purpose. Therefore, we formulate:

*Hypothesis 4: Borrowers exhibit preferences for minimal debt
distinctiveness, selecting one loan use*

Notwithstanding, credit choices relate to non-structural factors.
@cooper_consumption_2023 show that younger borrowers face constraints in
the types of loans they can access in typical credit markets. The
authors argue that this constraint shapes how these borrowers form loan
portfolios, often concentrating around different use cases such as
education, personal consumption, or small business loans. Considering
the systemic differences in how age affects credit consumption, we
assert that younger borrowers have different loan use portfolios. The
ERGM term `b1cov("age")` can capture this by assessing how changes in
age affect how borrowers form connections with different loan types.
Therefore:

*Hypothesis 5: Younger borrowers have different loan use mixes than
older borrowers*

Alongside age, the borrower's gender is known to be a significant factor
in influencing credit use decisions. @aliano_role_2023 study the role of
gender and other factors on borrowers' probability of default on the
Bondora.com, finding a persistent gender effect on different loan uses;
women tend to default less on health, home, and business loans.
@croson_gender_2009 shows that these differences can be due to risk
perception and self-selection effects, where women exhibit greater risk
aversion, leading to different borrowing patterns. We expect that these
differences in risk tolerance and perceived creditworthiness by lenders
affect loan use composition in terms of gender. The term
`b1nodematch("gender")` captures this effect by assessing homophily in
loan use selection. Our premise is supported by a negative estimate for
the term, indicating heterophily. Therefore:

*Hypothesis 6: Gender differences lead to different loan use mixes*

| Hypothesis | Term | Explanation |
|---------------------|-----------------|----------------------------------|
| $H_3^a:$  Borrowers who share one loan purpose are likely to share another | *cycle(4)* | This captures the tendency of borrowers and loan uses to be cyclical, where people cluster around shared uses of loans. |
| $H_4^a:$ Borrowers exhibit preferences for minimal debt distinctiveness, selecting one loan use | *b1degree(1)* | This captures the tendency of borrowers to prefer minimal distinctiveness in their loan uses, preferring to hold only one distinct type to minimise their perceived risk. |
| $H_5^a:$ Younger borrowers have different loan use mixes than older borrowers | *b1cov("age")* | This captures the tendency for younger borrowers to choose different kinds of loans based on how differently they consume credit. |
| $H_6^a:$ Gender differences lead to different loan use mixes | *b1nodematch("gender")* | This captures the idea that there are systemic differences in what kinds of loan uses are preferred the genders due to their perceived riskiness and creditworthiness. |

: Hypotheses related to Study 2

Following this section, we outline the methodology, explaining our
dataset and network construction. To facilitate this, descriptive
statistics are analysed for both the dataset and resulting network.
Subsequently, we elaborate on our choice of models, discussing how
network models can be used to conduct the study. Then, modelling results
are shown and explained to evaluate our hypotheses. Simultaneously, the
robustness of the study is evaluated through the models' diagnostics.
Finally, we arrive at our conclusions, summarising the study, its
results, and implications. Supporting items such as the source code and
specific data processing steps are shown in section \ref{sec:supp}.

## Dataset {#sec-methodology}

The study utilises publicly-available data from Bondora, a European P2P
lending platform primarily operating in the Baltics and Spain. The
dataset contains detailed information on defaulted and non-defaulted
loans granted to users between February 2009 and July 2021.
Specifically, contained is a range of numeric, binary, categorical, and
time-series attributes across 85,087 unique users and 179,235 individual
loans. Since the company's API is no longer accessible, we utilised a
publicly available repository compiled by @manu_siddhartha_bondora_2021
on kaggle.com. The user made a series of API calls to collect the data.
Following this, the dataset was processed according to Appendix
\ref{sec:preprocessing}.

### Pre-Processing & Network Formation

Initially, the dataset was filtered loans between 2014 and 2016 since
the data is incomplete during other periods. Since we are studying
interactions across two distinct set of nodes, a bipartite network was
formed with unique users as the first partition and the purpose of loans
as the second. This allows us to evaluate how borrowers interact with
loan purposes; a one-mode projection onto users is not meaningful for
P2P networks as it produces no discernible structure. For model
efficiency and performance, we randomly sample 500 individuals from the
larger sample. Ultimately, we have 500 and 9 nodes in the first and
second partitions, respectively.

### Descriptive Statistics & Preliminary Analysis

@tbl-desc-table summarises the numeric variables between the reduced and
larger samples. First, borrowers take out relatively small loans with
fairly high interest rates, possibly reflecting that the average
borrower is deemed fairly risky. The average loan is quite long, at
approximately 46 months out of a maximum 60. @fig-bar-rating shows that
the largest credit category is HR[^1], which is the lowest possible
rating. So, the typical borrower is relatively uncreditworthy. Moreover,
the largest categories of loan use are "Other" and "Home Improvement",
suggesting these are largely personal. We observe minimal differences
between the random sampling and the larger sample of individuals, per
@tbl-desc-table and @fig-bar-rating, however, the study may be biased by
the fact that the dataset contains only granted loans, rather than all
bids for loans. This means our inferences about how borrowers behave is
limited to successful bids. Further, the time period is relatively
distant, meaning that recent structural changes to the P2P sector cannot
be accounted for.

[^1]: Bondora uses their own proprietary credit rating system where the
    best rating is AA and worst is HR [@bondora_how_2024]. The company
    bases these ratings on their calculated expected probability of loss
    on the loan. For example, someone rated AA ranges from a expected
    loss of 0% to 2%, whereas for HR this is 25% - \>25%.

```{r}
#| label: tbl-desc-table
#| tbl-cap: "Descriptive Statistics of Numeric Variables"

knitr::kable(sum_tab)
```

@tbl-net-stats shows a network density of 0.12, indicating that only 12%
of all possible borrower–loan-type connections exist. This density
suggests that borrowers typically engage with few loan types, reflecting
specialization in borrowing behaviour. The centralization value of 0.47
further indicates an uneven distribution across loan types, where a few
popular categories attract many borrowers while most remain peripheral.
The mean distance of approximately three implies that borrowers are
closely connected, typically separated by only three-to-four loan-use
steps.

Centrality distributions reinforce this structure [@fig-network-plots].
Betweenness is highest among loan-type nodes, though several borrowers
also bridge distinct loan clusters. Closeness centrality remains low but
scattered, suggesting localized clustering constrained by the bipartite
structure. Degree distributions show wide variation among loan types,
consistent with a hub-like structure dominated by a few popular uses.
However, the second partition has relatively small degree dispersion.
Finally, the homogeneous eccentricity distribution indicates that no
borrowers are highly isolated, reflecting a compact and moderately
cohesive network overall.

```{r}
#| label: tbl-net-stats
#| tbl-cap: "Basic Network Descriptive Statistics"

knitr::kable(readRDS(here::here("resources","objects","ergm","net_summary.Rds")))
```

```{r fig.width=14, fig.height=7}
#| label: fig-network
#| fig-cap: "Network Visualisation"

replayPlot(net_plot)
```

\begin{footnotesize}
\linespread{1}
Note: Each colour represents a different partition of the network. The partitions' nodes are scaled to their relative degrees. Users with many different loan types are larger than those with single loan uses. Similarly, the size of second partition nodes represents how many users belong to each.
\end{footnotesize}

```{r fig.width=14, fig.height=7}
#| label: fig-network-plots
#| fig-cap: "Network Centrality Distribution Plots"

replayPlot(cent_plots)
```

## Research Rationale

Our overall study analyses borrowers' behaviours, however, the studies
are differentiated in their scope; the first study observes dyad-level
relations among borrowers to understand how shared attributes affect
borrowers' behaviours. This is inherently useful because behavioural
finance aims to understand how individuals behave in *aggregate*
[@ayal_behavioral_2018]. Typically, the Quadratic Assignment Procedure
is selected for dyad-level observations because basic linear regressions
such as Ordinary Least Squares are fundamentally flawed and biased when
using non-independent observations [@simpson_qap_2001].

Recognising this, studies such as @zuo_clustering_nodate assume that
groups of borrowers have dependencies, and accordingly, use clustering
methods such as K-Means or Fuzzy Clustering to understand behaviour.
However, these methods are unsuitable when we require a distribution of
estimates to make inferences from our results. Moreover, studying
individuals as dyads rather than large clusters provides finer
granularity in results. Therefore, QAP regression is the most
appropriate choice for this study: it allows us to control for the
non-independence among dyads, derive a statistical distribution of
estimates through matrix permutation, and draw valid inferences about
the relationship between borrower similarities and loan-use behaviour.
This makes it both theoretically and methodologically well-suited to our
research objectives and dataset.

The second study extends the analysis from a dyadic similarities to the
structural formation of borrower-to-loan-use connections, which can
identify the behavioural mechanisms that influence borrowers' specific
loan use mixes. While QAP can account for exogenous influences, it
cannot analyse higher-order network patterns such as clustering while
accounting for simultaneous interactions between borrowers and loan
uses. Using the Exponential Random Graph Model allows us to analyse how
local behavioural tendencies aggregates to a global network structure.

Recent evidence highlights the need to focus on structural approaches in
P2P lending. @liu_network_2024 studies borrowers' centrality with the
Bondora P2P dataset, modelling edges through borrower-to-borrower
similarity metrics. Ultimately, they find that a borrower's position is
highly significant for their probability of default, using centrality as
an exogenous variable in a logistic regression. In contrast, ERGM allows
us to move away from predictive association and instead account for
endogenous factors, allowing us to understand the likelihood that
borrowers' selection of loan uses are random. Consequently, we can
better understand how behavioural phenomena such as familiarity and debt
account aversion can translate into *systemic* patterns of credit use.

As we aim to understand how two distinct groups of nodes interact, a
bipartite approach is highly suitable. Specifically,
@stivala_improving_2025 shows that in bipartite networks, terms such as
`cycle(4)` can be effective in capturing clustering as in hypothesis 3,
whereas it would fail in one-mode projected networks. Furthermore, the
flexibility of ERGM also allows for modelling exogenous impacts such as
heterophily in gender-based loan use mixes, as in hypothesis 6, and
covariate age effects as in hypothesis 5.

\clearpage
\appendix

## Supplements

\label{sec:supp}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\lstset{ %
  language=R,                     % the language of the code
  basicstyle=\footnotesize,       % the size of the fonts that are used for the code
  numbers=left,                   % where to put the line-numbers
  numberstyle=\tiny\color{gray},  % the style that is used for the line-numbers
  stepnumber=1,                   % the step between two line-numbers. If it's 1, each line
                                  % will be numbered
  numbersep=5pt,                  % how far the line-numbers are from the code
  backgroundcolor=\color{white},  % choose the background color. You must add 
  showspaces=false,               % show spaces adding particular underscores
  showstringspaces=false,         % underline spaces within strings
  showtabs=false,                 % show tabs within strings adding particular underscores
  frame=false,                   % adds a frame around the code
  rulecolor=\color{black},        % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. commens (green here))
  tabsize=2,                      % sets default tabsize to 2 spaces
  captionpos=b,                   % sets the caption-position to bottom
  breaklines=true,                % sets automatic line breaking
  breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
  title=\lstname,                 % show the filename of files included with \lstinputlisting;
                                  % also try caption instead of title
  keywordstyle=\color{blue},      % keyword style
  commentstyle=\color{dkgreen},   % comment style
  stringstyle=\color{mauve},      % string literal style
}

### Data Preprocessing Steps

\label{sec:preprocessing}

This section outlines the specific steps taken to process the data into
the network object that was used in later analysis, outlining the
rationale, where needed.

-   Following the data import, only the following attributes were kept
    to make our processing steps more focused

``` r
keep_cols <- c("LoanId", "UserName","Age", "Gender", 
               "Country", "Amount", "Interest","LoanDuration",
              "UseOfLoan", "Rating", "Restructured", "MonthlyPayment",
              "OccupationArea", "BiddingStartedOn")
```

-   We removed rows with any `NA` values for the selected attributes to
    ensure that we have a complete dataset

-   Observing @fig-user-counts, we see that between 2017-2021, there
    exists many -1 values for `UseofLoan`. This is because Bondora no
    longer provided this detail in later datasets. However, between 2014
    and 2016 this data is available. Because of this, we restrict the
    time period to have the least noisy data. `BiddingStartedOn` was
    chosen to record the time, because the auction start date can be a
    good indication of activity around the loan.

```{r fig.width=14, fig.height=5}
#| label: fig-user-counts
#| fig-cap: "Number of Unique Users per Loan Type per Year"

replayPlot(readRDS(here::here("resourceS","objects","preprocessing","user_cnt_plt.Rds")))
```

-   Then, we sampled 500 individuals out of the filtered dataset to
    ensure that our models run sufficiently well and converge. By
    choosing 500 individuals, we preserve importance structures within
    the network while adding sufficient statistical power.

<!-- -->

-   When adding labels to each loan use and occupation area, we keep any
    possible missing values because these do not indicate lower quality
    data. Instead, they can carry information about lenders distribute
    loans according to the quality of information presented by borrowers
    [@yao_relationship_2019]. Additionally, keeping missing labels is
    important to ensure that the edges are not superficially limited and
    that nodes can be isolates if they are indeed that way in reality.

-   To create the network object for ERGM models:

    -   We first create an incidence matrix of size $n \times m$,
        borrower usernames by the loan use, where the values are the
        number of loans the user obtained for each loan use category.
        These values are then turned into binary because our modelling
        does not support the use of weighted edges.

    -   A bipartite network is created from this incidence matrix, where
        the first partition are unique users, and the second partition
        is the loan use associated among all loans of the user. Users
        can share more than more loan use type.

    -   We add age and gender attributes to each node of the first
        partition

-   To create the relevant networks/matrices for QAP Regressions:

    -   We first re-create the loan use incidence matrix, however, we
        keep the counts as the linear regression supports this.

        -   We then convert the loan use incidence matrix into an
            adjacency matrix of shape $n\times n$ through the
            transformation $\mathbf{X}\cdot\mathbf{X^T}$, which results
            in a matrix of shared loan uses weighted by the frequency of
            each user's loan count within each loan use category.

        -   We then set the diagonal values of $\mathbf{X}$ to be zero
            to ensure that there are no loops within the network.

    -   In accordance with this procedure, we create adjacency matrices
        for Credit Rating and Occupation Area which will be used as the
        main predictors in the QAP models.

        -   Credit Rating is a weighted adjacency matrix based on the
            loan count belonging to each user for each credit rating
            category

        -   Occupation Area is a binary adjacency matrix based on which
            occupation each user has reported to have held in the past
            when applying for each of their loans.

    -   Additionally, we create a set of control variables to improve
        the validity and performance of the linear regression using Loan
        Amounts, Age, Gender, Loan Duration, and whether the loans have
        been restructured.

        -   Loan Amounts is a weighted adjacency matrix based on five
            bins across the range of possible loan amounts

        -   Age is an adjacency matrix based on differences in users'
            ages

        -   Gender is a binary adjacency matrix based on users' shared
            gender.

        -   Loan Duration is an adjacency matrix based on the
            differences in users' loans average durations

        -   Restructured is a binary adjacency matrix based on whether
            the users share the fact that they have defaulted on any
            loans in the past. The users can either have shared both
            defaults and non-defaults, or both.

\clearpage

### Distributions of Variables across Samples

\label{sec:dist}

```{r fig.width=14, fig.height=5}
#| label: fig-hist-amt
#| fig-cap: "Distribution of Loan Amount across Samples"
#| warning: false

replayPlot(readRDS(here::here("resources","objects","preprocessing",
                          "hist_amt.Rds")))
```

```{r fig.width=14, fig.height=5}
#| label: fig-hist-int
#| fig-cap: "Distribution of Interest Rates across Samples"
#| warning: false

replayPlot(readRDS(here::here("resources","objects","preprocessing",
                          "hist_int.Rds")))
```

```{r fig.width=14, fig.height=5}
#| label: fig-hist-age
#| fig-cap: "Distribution of Age across Samples"
#| warning: false

replayPlot(readRDS(here::here("resources","objects","preprocessing",
                          "hist_age.Rds")))
```

```{r fig.width=14, fig.height=5}
#| label: fig-hist-loandur
#| fig-cap: "Distribution of Loan Duration across Samples"
#| warning: false

replayPlot(readRDS(here::here("resources","objects","preprocessing",
                          "hist_loandur.Rds")))
```

```{r fig.width=14, fig.height=5}
#| label: fig-bar-loanuse
#| fig-cap: "Distribution of Loan Purpose across Samples"
#| warning: false

replayPlot(readRDS(here::here("resources","objects","preprocessing",
                          "bar_loanuse.Rds")))
```

```{r fig.width=14, fig.height=5}
#| label: fig-bar-rating
#| fig-cap: "Distribution of Credit Ratings across Samples"
#| warning: false

replayPlot(readRDS(here::here("resources","objects","preprocessing",
                          "bar_rating.Rds")))
```

```{r fig.width=14, fig.height=7}
#| label: fig-bar-occupation
#| fig-cap: "Distribution of Occupation Types across Samples"
#| warning: false

replayPlot(readRDS(here::here("resources","objects","preprocessing",
                          "bar_occupation.Rds")))
```

\clearpage

### Source Code - Data Preprocessing

\linespread{1} 
\lstinputlisting[language=R]{scripts/bondora_preprocessing.R}
\clearpage

### Source Code - QAP Linear Regression

\linespread{1} 
\lstinputlisting[language=R]{scripts/qap_network_analysis.R} 
\clearpage

### Source Code - ERGM Network Analysis

\linespread{1} 
\lstinputlisting[language=R]{scripts/ergm_network_analysis.R} 
\clearpage
\linespread{1.15}

## References {.unnumbered}

::: {#refs}
:::

\clearpage

## Technology Statement

> During the preparation of this work, we used ChatGPT in order to
> generate select parts of the R script utilised to process the dataset.
> Specifically, the tool was used to transform the processed dataset
> into a format that `igraph` would accept as a network object.
> Additionally, some R functions written by the group were improved with
> ChatGPT's suggestions. No AI tool was utilised to independently write
> parts of the report. The following parts of the assignment were
> affected/generated by AI tool usage: **INTRODUCTION;** The tool was
> utilised to evaluate the validity of certain theoretical concepts and
> refine them. **DATASET;** the data described within this section was
> processed partly by some code drafted by ChatGPT and edited by the
> group. After using this tool/service, **Group 7** evaluated the
> validity of the tool’s outputs, including the sources that generative
> AI tools have used, and edited the content as needed. As a
> consequence, **Group 7** takes full responsibility for the content of
> their work.
