---
title: "Credit Allocation Mix in Peer-to-Peer Lending: A Network Study on Bondora"
subtitle: "Social Network Analysis - Group 7"
date: 2025-12-14
author:
  - name: Floris 	Vermeulen
    affiliations:
      - ref: jads
      - ref: tiu
  - name: Martijn van Iterson
    affiliations:
      - ref: jads
      - ref: tiu
  - name: Niek	Fleerakkers
    affiliations:
      - ref: jads
      - ref: tiu
  - name: Patryk	Grodek
    affiliations:
      - ref: jads
      - ref: tue
  - name: Samir Sabitli
    affiliations:
      - ref: jads
      - ref: tue
affiliations:
  - id: jads
    name: Jheronimus Academy of Data Science
  - id: tiu
    name: Tilburg University
  - id: tue
    name: Eindhoven University of Technology
include-in-header:
  - text: |
      \usepackage{tgschola}
      \usepackage[T1]{fontenc}
      
      \usepackage{longtable}
      \usepackage{graphicx}
      \usepackage{booktabs}
      \usepackage{dcolumn}
      \usepackage{float}
      \usepackage{listings}
      \usepackage{color}
      \linespread{1.15}
      \usepackage{sectsty}
      \allsectionsfont{\sffamily}
bibliography: resources/r-references.bib
csl: resources/apa-6th-edition.csl
toc: true
link-citations: true
toc-title: "Table of Contents"
toc-depth: 4
format:
  pdf: 
    pdf-engine: xelatex
    fig-pos: H
    lof: true
    lot: true
    documentclass: article
    pagesize: a4
    number-sections: true
    df-print: kable
    fig-cap-location: top
    mainfont: "TeX Gyre Schola"
    mathfont: "Latin Modern Math"
    sansfont: "Arial"
    geometry:
      - left=1.15in
      - right=1.15in
      - bottom=1.15in
      - top=1.15in
      - heightrounded
execute:
    echo: false
    cache: false
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| label: analysis-preferences
#| include: false

# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
#install.packages("here")
#install.packages("gridExtra")
```

```{r}
# Include any relevant objects
net <- readRDS(here::here("resources","objects","preprocessing",
                          "bondora_net.Rds"))
net_plot <- readRDS(here::here("resources","objects","ergm",
                          "network_plot.Rds"))
sum_tab <- readRDS(here::here("resources","objects","preprocessing",
                          "summary_table.Rds"))
cent_plots <- readRDS(here::here("resources","objects","ergm",
                          "network_plots.Rds"))
```

{{< pagebreak >}}

\renewcommand{\thesection}{\Roman{section}}
\pagenumbering{roman}

## Abstract

In this study we investigate the behavioural phenomena that influence
individual borrowers' choice of loan purposes. In the first study, we
evaluate how loan counts within each loan purpose choice are influenced
by borrowers' individual characteristics by asking "*Do borrowers vary
in their loan purpose type and frequency based on the individual and
loan characteristics?".* Using the Quadratic Assignment Procedure (QAP)
linear regression framework. Our first hypothesis asserts that borrowers
sharing the same credit rating should choose similar loan purposes, but
with a preference for a lower loan count. We found that while this is a
significant effect, the impact on loan count is minimal. The second
asserts that borrowers sharing the same occupation area should share
loan purposes with a lower count as well. Here, we do not find a
meaningful effect. The second study examines which behavioural phenomena
affect how borrowers decide which combination of loan purposes to
request. We ask, *"What behavioural mechanisms affect the structure of
borrower-to-loan-use connections"*. With the use of Exponential Random
Graph Models (ERGM), we first determine whether there is co-affiliation
between borrowers who overlap in their loan choices. We find that there
is no systematic aggregate overlapping effect. Further, we evaluate
whether there is an individual behavioural effect where borrowers
minimise the number of loan types. The results indicate a systematic
individual behavioural bias. Assessing whether younger borrowers have
different loan purpose selections, we fail to find evidence for this
assertion. Similarly, we fail to find support for the idea that there
are gender differences in how borrowers choose loan types.

\clearpage
\pagenumbering{arabic}
\setcounter{section}{0}
\renewcommand{\thesection}{\arabic{section}}

## Introduction {#sec-intro}

With the advent of online peer-to-peer (P2P) lending platforms, the
traditional methods of financial intermediation have been usurped by
individual choice; with no stringent third parties involved in
transactions, individuals have greater power in sourcing credit. This
presents new realities for individuals disenfranchised by the
traditional financial system [@alistair_milne_business_2017].
Traditional financial literature shows that individually, investors
exhibit behavioural biases [@agrawal_managing_2012; @ayal_two_2011] in
how they choose their investments. However, numerous studies demonstrate
that these behavioural biases are present among non-professional P2P
borrowers. For example, @herzenstein_strategic_2011 and
@lee_herding_2012 suggest that users cluster around popular loans,
exhibiting 'herding' behaviour. Literature focuses predominantly on
funding success [@yao_relationship_2019], however, @ayal_behavioral_2018
suggests that there is insufficient attention on how borrowers form
their loan portfolios. As the authors indicate, this area of research is
important in understanding why non-professional investors deviate from
known models of rational investor behaviour. Our study aims to model and
understand how P2P borrowers' behaviours influence what kinds of loans
they acquire. In subsequent discussions, we consider investors as
rational participants in credit markets, as described by theory, whereas
borrowers refer to the users of P2P platforms.

Investigating behavioural biases, @ayal_behavioral_2018 find that
investors' familiarity with specific assets can make them seem less
risky, making recognisable assets more attractive. However, such assets
lead to undiversified portfolios. Evidence with P2P platforms supports
the familiarity bias in lenders, as @galak_micro-finance_2010
demonstrate that lenders choose borrowers who are similar across
demographic attributes such as gender, occupation, and ethnicity.
Moreover, @amar_winning_2011 explain that borrowers show an aversion to
holding a *large number of debts,* preferring to first reduce debt
count, even if they had other debts with higher interest rates. This
irrational behaviour demonstrates that individual borrowers face
behavioural biases on the *type* and *number* of debts they select.
Based on these notions, we formulate our first study, asking:

***Research Question 1:** Do borrowers vary in their loan purpose type
and frequency based on their individual and loan characteristics?*

This question is answered using linear Quadratic Assignment Procedure
(QAP) regressions, as we can represent similarities across loan purposes
and attributes as matrices, weighted by loan purpose frequency, as
inputs in the model.

Using data from LendingClub, @serrano-cinca_determinants_2015 study
funding success, finding that the reported loan purpose is a significant
factor explaining differences in default rates. Specifically, the
authors argue that there is heterogeneity in the credit riskiness of
individuals seeking specific loan purposes, indicating that credit
rating can explain some variation in borrowers' choice of loan purpose.
Considering the preference for lower loan counts as well, we formulate
the following hypothesis:

*Hypothesis 1: Borrowers with the same credit ratings choose fewer loans
within their shared loan purpose*

Supporting the familiarity bias, @ravina_love_2019 studied personal
characteristics in P2P markets, finding that loans tend to be awarded to
those sharing personal characteristics such as occupation. As
@ayal_behavioral_2018 imply, this can be generalised to suggest that
lenders' loan grants reflect a wider network of similar borrower
preferences, which can reflect the homogeneity in loan purpose observed
by @serrano-cinca_determinants_2015. Therefore, we expect occupation to
be a meaningful predictor. However, per @amar_winning_2011, we expect a
preference for smaller loan counts within borrowers' shared occupations.
Thus,

*Hypothesis 2: Borrowers sharing the same occupation area tend to share
fewer loans for their shared loan purpose.*

Further, behavioural finance suggests that individuals' financial
decisions are not independent, but rather, shaped by bounded rationality
and psychological preferences [@kahneman_prospect_1979]. In the P2P
context, @ayal_behavioral_2018 shows that these preferences can
influence how borrowers choose loan types and how this can overlap with
other borrowers' decisions. Using the ERGM framework, we can identify
how these behaviours manifest as structural regularities in a
borrower-to-loan-use network. Therefore, we ask:

***Research Question 2:** What behavioural mechanisms affect the
structure of borrower-to-loan-use connections?*

The previous section outlined that the familiarity bias can lead to
similarities across borrower portfolios in terms of their
characteristics. However, in P2P settings, this can manifest itself as
familiarity *clustering*, wherein borrowers who have previously been
granted certain loan types are more likely to repeat their choices, or
select other categories that appear familiar to different borrowers.
@herzenstein_strategic_2011, studying borrower decisions on Prosper.com,
identified a 'strategic herding' behaviour, where individuals, **in
aggregate,** gravitate towards popular or socially-validated loans.
Considering that loan popularity is directly observable through its
number of bids, we assert that that the social validation of a loan
purpose for one borrower positively reinforces the selection of that
same loan purpose for other, **co-affiliated** borrowers.

In network terms, this aggregate herding behaviour generates a
structural propensity for four-cycles to form, resulting in closed paths
between overlapping nodes of each of the bipartite partitions, as
suggested by @koskinen_modelling_2012. This structure represents the
fact that two borrowers connected to a common loan purpose have a higher
likelihood of also connecting to a *second* common loan purpose. We
capture this specific form of bipartite closure through the ERGM term
`cycle(4)`, formulating:

*Hypothesis 3: Co-affiliated borrowers are more likely to overlap on a
second loan purpose.*

Further, @ayal_perceived_2009, expand on the theory of Debt Account
Aversion with the Theory of Perceived Diversification, explaining that
**individual** borrowers' distress increases with not only with number
of debts, but also the perceived *distinctiveness* of each debt type,
suggesting that in the P2P context, debt that is mentally differentiated
by its purpose would also be seen as distressing. Therefore, we assert
that if borrowers do not diversify, then they try to minimise loan
distinctiveness by requesting loans of a single type. This can be
captured by the `b1degree(1)` term, which models whether borrowers are
only linked to one loan purpose. Therefore, we formulate:

*Hypothesis 4: Borrowers exhibit preferences for minimal debt
distinctiveness, selecting one loan purpose.*

Notwithstanding, credit choices relate to non-structural factors.
@cooper_consumption_2023 show that younger borrowers face constraints in
the types of loans they can access in typical credit markets. The
authors argue that this constraint shapes how these borrowers form loan
portfolios, often concentrating around different use cases such as
education, personal consumption, or small business loans. Considering
the systemic differences in how age affects credit consumption, we
assert that younger borrowers have different loan purpose portfolios.
The ERGM term `b1cov("age")` can capture this by assessing how changes
in age affect how borrowers form connections with different loan types.
Therefore:

*Hypothesis 5: Younger borrowers have different loan purpose mixes than
older ones.*

Alongside age, the borrower's gender is known to be a significant factor
in influencing credit use decisions. @aliano_role_2023 studied the role
of gender and other factors on borrowers' probability of default on
Bondora.com, finding a persistent gender effect on different loan
purposes; women tend to default less on health, home, and business
loans. @croson_gender_2009 show that these differences can be due to
risk perception and self-selection effects, where women exhibit greater
risk aversion, leading to different borrowing patterns. We expect that
these differences in risk tolerance and perceived creditworthiness by
lenders affect loan purpose composition in terms of gender. The term
`b1factor("gender")` captures this effect by assessing whether
borrowers' connections with loan purposes differ across genders. We
would find support for this idea given a significant, non-zero effect
estimate. Therefore:

*Hypothesis 6: Gender differences lead to different loan purpose mixes.*

| Hypothesis | Term | Explanation |
|----------------------|--------------|----------------------------------|
| $H_3^a:$ Co-affiliated borrowers are more likely to overlap on a second loan purpose | *cycle(4)* | This captures the tendency of borrowers who are already affiliated through their connections to the same loan type to have connections to another type. With this term we capture **aggregate** borrower behaviour. |
| $H_4^a:$ Borrowers exhibit preferences for minimal debt distinctiveness, selecting one loan purpose | *b1degree(1)* | This captures the tendency of borrowers to prefer minimal distinctiveness in their loan purposes, preferring to hold only one distinct type to minimise their perceived risk. With this term, we capture **individual** borrower behaviour rather than an aggregation phenomenon. |
| $H_5^a:$ Younger borrowers have different loan purpose mixes than older borrowers | *b1cov("age")* | This captures the tendency for younger borrowers to choose different kinds of loans based on how differently they consume credit. |
| $H_6^a:$ Gender differences lead to different loan purpose mixes | *b1factor("gender")* | There are systemic differences in what kinds of loan purposes are preferred the genders due to their perceived riskiness and creditworthiness. |

: Hypotheses related to Study 2

Following this section, we outline the methodology, explaining our
dataset and network construction. To facilitate this, descriptive
statistics are analysed for both the dataset and resulting network.
Subsequently, we elaborate on our choice of models, discussing how
network models can be used to conduct the study. Then, modelling results
are shown and explained to evaluate our hypotheses. Simultaneously, the
robustness of the study is evaluated through the models' diagnostics.
Finally, we arrive at our conclusions, summarising the study, its
results, and implications. Supporting items such as the source code and
specific data processing steps are shown in section \ref{sec:supp}.

## Dataset {#sec-methodology}

The study utilises publicly-available data from Bondora, a European P2P
lending platform, primarily operating in the Baltics and Spain. The
dataset contains detailed information on defaulted and non-defaulted
loans granted to users between February 2009 and July 2021.
Specifically, contained is a range of numeric, binary, categorical, and
time-series attributes across `85,087` unique users and `179,235`
individual loans. Since the company's API is no longer accessible, we
utilised a publicly available repository compiled by
@manu_siddhartha_bondora_2021 on kaggle.com. The data compiler made a
series of API calls to collect the data. Following this, the dataset was
processed according to Appendix Section \ref{sec:preprocessing}.

### Pre-Processing & Network Formation

Initially, the dataset was filtered to loans between 2014 and 2016 since
the data is incomplete during other periods. Since we are studying
interactions across two distinct set of nodes, a bipartite network was
formed with unique users as the first partition and the purpose of loans
as the second. This allows us to evaluate how borrowers interact with
loan purposes; a one-mode projection onto users is not meaningful for
P2P networks as it produces no discernible structure. For model
efficiency and performance, we randomly sampled `500` individuals from
the larger sample. Ultimately, we have `500` nodes of users and `9`
nodes of loan purposes, in the first and second partitions,
respectively.

### Descriptive Statistics & Preliminary Analysis

@tbl-desc-table summarises the numeric variables for the reduced and
larger samples. First, borrowers take out relatively small loans with
fairly high interest rates, possibly reflecting that the average
borrower is deemed fairly risky. The average loan is quite long, at
approximately `46` months out of a maximum of `60`. @fig-bar-rating
shows that the largest credit category is `HR`[^1], which is the lowest
possible rating. So, the typical borrower is rated as relatively
uncreditworthy. Moreover, the largest categories of loan purpose are
`Other` and `Home Improvement`, suggesting these are largely personal.
Additionally, @fig-bar-country shows that the biggest borrower group is
from Estonia. However, Spain and Finland are also highly represented.
Conversely, Slovakia has minimal representation. We observe minimal
differences between the random sampling and the larger sample of
individuals, per @tbl-desc-table and @fig-bar-rating. However, the study
may be biased by the fact that the dataset contains only granted loans,
rather than all bids for loans. This means our inferences about how
borrowers behave is limited to successful bids. Further, the time period
is relatively distant, meaning that recent structural changes to the P2P
sector cannot be accounted for.

[^1]: Bondora uses their own proprietary credit rating system where the
    best rating is AA and worst is HR [@bondora_how_2024]. The company
    bases these ratings on their calculated expected probability of loss
    on the loan. For example, someone rated AA ranges from a expected
    loss of 0% to 2%, whereas for HR this is 25% - \>25%.

```{r}
#| label: tbl-desc-table
#| tbl-cap: "Descriptive Statistics of Numeric Variables"

knitr::kable(sum_tab)
```

@tbl-net-stats shows a network density of `0.12`, indicating that only
12% of all possible borrower–loan-type connections exist. This density
suggests that borrowers typically engage with few loan types, reflecting
specialization in borrowing behaviour. The centralization value of
`0.47` further indicates an uneven distribution across loan types, where
a few popular categories attract many borrowers while most remain
peripheral. The mean distance of approximately three implies that
borrowers are closely connected, typically separated by only
three-to-four loan-use steps.

Centrality distributions reinforce this structure as shown in
[@fig-network-plots]. Betweenness is highest among loan-type nodes,
though several borrowers also bridge distinct loan clusters. Closeness
centrality remains low but scattered, suggesting localised clustering
constrained by the bipartite structure. Degree distributions show wide
variation among loan types, consistent with a hub-like structure
dominated by a few popular uses. However, the second partition has
relatively small degree dispersion. Finally, the homogeneous
eccentricity distribution indicates that no borrowers are highly
isolated, reflecting a compact and moderately cohesive network overall.

```{r}
#| label: tbl-net-stats
#| tbl-cap: "Basic Network Descriptive Statistics"

knitr::kable(readRDS(here::here("resources","objects","ergm","net_summary.Rds")))
```

```{r fig.width=14, fig.height=7}
#| label: fig-network
#| fig-cap: "Network Visualisation"

replayPlot(net_plot)
```

\begin{footnotesize}
\linespread{1}
Note: Each colour represents a different partition of the network. The partitions' nodes are scaled to their relative degrees. Users with many different loan types are larger than those with single loan purposes. Similarly, the size of second partition nodes represents how many users belong to each.
\end{footnotesize}

```{r fig.width=14, fig.height=10}
#| label: fig-network-plots
#| fig-cap: "Network Centrality Distribution Plots"

replayPlot(cent_plots)
```

## Research Rationale

We investigate borrower behaviour, however, the studies are
differentiated in their scope; the first study observes dyad-level
relations among borrowers to understand how shared attributes affect
borrowers' behaviour. This is inherently useful because behavioural
finance aims to understand how individuals behave in *aggregate*
[@ayal_behavioral_2018]. Typically, the Quadratic Assignment Procedure
is selected for dyad-level observations because basic linear regressions
such as Ordinary Least Squares are fundamentally flawed and biased when
using non-independent observations [@simpson_qap_2001].

Recognising this, studies such as @zuo_clustering_2015 assume that
groups of borrowers have dependencies, and accordingly, use clustering
methods such as K-Means or Fuzzy Clustering to understand behaviour.
However, these methods are unsuitable when we require a distribution of
estimates to make inferences from our results. Moreover, studying
individuals as dyads rather than large clusters provides finer
granularity in results. Therefore, QAP regression is the most
appropriate choice for this study: it allows us to control for the
non-independence among dyads, derive a statistical distribution of
estimates through matrix permutation, and draw valid inferences about
the relationship between borrower similarities and loan-use behaviour.
This makes it both theoretically and methodologically well-suited to our
research objectives and dataset.

The second study extends the analysis from dyadic similarities to the
structural formation of borrower-to-loan-use connections, which can
identify the behavioural mechanisms that influence borrowers' specific
loan purpose mixes. While QAP can account for exogenous influences, it
cannot analyse higher-order network patterns such as clustering while
accounting for simultaneous interactions between borrowers and loan
purposes. Using ERGM allows us to analyse how local behavioural
tendencies aggregate to a global network structure.

Recent evidence highlights the need to focus on structural approaches in
P2P lending. @liu_network_2024 studied borrowers' centrality with the
Bondora P2P dataset, modelling edges through borrower-to-borrower
similarity metrics. Ultimately, they find that a borrower's position is
highly significant for their probability of default, using centrality as
an exogenous variable in a logistic regression. In contrast, ERGM allows
us to move away from predictive association and instead account for
endogenous factors, allowing us to understand the likelihood that
borrowers' selections of loan purposes is random. Consequently, we can
better understand how behavioural phenomena such as familiarity and debt
account aversion can translate into *systemic* patterns of credit use.

As we aim to understand how two distinct groups of nodes interact, a
bipartite approach is highly suitable. Specifically,
@koskinen_modelling_2012 shows that in bipartite networks, terms such as
`cycle(4)` can be effective in capturing the idea of grouping across
partitions, as in hypothesis 3, whereas it would fail in a
fully-connected one-mode projected networks. Furthermore, the
flexibility of ERGM also allows for modelling exogenous impacts such
gender-based differences in loan purpose mixes, as in hypothesis 6, and
covariate age effects as in hypothesis 5.

## Results

### Study 1 - QAP Linear Regression

The first study is assessed with two QAP linear regression models, as
shown in @fig-qap-coef and Table \ref{tab-qap-results}. The first model
includes only the main predictors whereas the second controls for
borrowers' age, gender, loan amount, loan duration, and whether the loan
was restructured. Differences across these models can indicate the
strength of omitted variable bias.

Assessing hypothesis 1, in the unrestricted model, borrowers sharing the
same credit rating once leads to a `0.083` increase in the number of
loans they share within the same loan purpose category, holding all
other effects fixed. While this effect is statistically significant at
1%, notably, it is not economically significant because the effect
magnitude is very small. In essence, having the same credit rating as
another borrower does not mean that the borrower will have meaningfully
different amounts of loans taken out for the same reason. Without
controls, this effect was slightly larger, meaning that other predictors
capture some of this effect. Our results corroborate aspects of
@serrano-cinca_determinants_2015; similarly, we find that there is a
non-zero effect of creditworthiness on loan purpose, however, they find
a stronger effect. This can be due to borrowers' aversion to holding
large numbers of debt, as explained by @ayal_two_2011, where borrowers
attempt to minimize debt counts independently of the loan purpose type.

```{r fig.width=10, fig.height=5}
#| label: fig-qap-coef
#| fig-cap: "QAP Coefficient Plots"

p1 <- readRDS(here::here("resources","objects","qap","qap_m1_coef_plot.Rds"))
p2 <- readRDS(here::here("resources","objects","qap","qap_m3_coef_plot.Rds"))

gridExtra::grid.arrange(p1, p2, ncol = 2)
```

Further, assessing hypothesis 2, borrowers sharing the same occupation
area does not have a significant impact on the outcome. This indicates
that the number of loans taken out for the same reason as another
borrower does not vary, given that pairs of borrowers have the same kind
of job. The omission of control variables does not change this effect,
meaning that its insignificance is not due to other correlated effects.
The result is contrary to @ravina_love_2019, who found that certain
occupation areas have a higher likelihood of loan success. As the author
states, these effects are likely present in markets with sophisticated
decision makers. It is possible that Bondora lenders do not conform to
this standard. Alternatively, the familiarity bias lenders exhibit on
shared personal characteristics may not generalise to similarities
across borrower characteristics, as we have argued.

Observing the controls, we see that loan amounts and whether the loan
were restructured are statistically significant at 5%, but not
economically meaningful. Moreover, the model diagnostics show an
$R^2=0.035$, suggesting that the linear QAP only explains 3.5% of the
variation observed in the data. This is rather small, indicating that
the linear model is not a good fit for this specific study. We argue
that borrowers are not biased against large debt counts per loan purpose
type as *dyads*, but rather, as part of a more complex network
structure. This is indicated by the low $R^2$, showing that this dyadic
relationship cannot be captured as a linear effect.

```{r}
#| output: asis
qap_table_items <- readRDS(here::here("resources","objects","qap",                                 "qap_tables.Rds"))

texreg::texreg(qap_table_items[[1]], 
               custom.model.names = qap_table_items[[2]],
               digits = 3,
               caption.above = TRUE,
               caption = "QAP Modelling Results",
               label = "tab-qap-results",
               booktabs = TRUE,
               float.pos = "!h",
               use.packages = FALSE,
               scalebox = 1,
               dcolumn = TRUE)
```

### Study 2 - ERGM

The second study is assessed with four iterative models shown in
@fig-ergm-coef and Table \ref{tab-ergm-results}; the fourth is
considered to be the full, unrestricted model. In each iteration,
additional terms are added and diagnosed in Appendix Section
\ref{sec:ergm_diags} by the Goodness of Fit (GOF) and Markov chain Monte
Carlo (MCMC) simulations. We first interpret model estimates, their
relation to our hypotheses, and conclude by evaluating the validity of
the fitted models.

In Model 4, the inclusion of `b1factor("gender")` has pushed the `edges`
term towards zero, making it insignificant. This suggests that the
combination of individual borrowers' loan purpose formation is likely
random in the presence of other terms, indicating that the edge
formation effect is instead captured jointly by the rest of the terms.
We consider this a good result since it suggests that our final model
may be fully saturated.

Evaluating hypothesis 4, we determine whether borrowers exhibit a
preference for minimal debt distinctiveness by choosing only one loan
type. In Model 4, `b1degree(1)` is highly significant and positive with
a log-odds of `3.271`, suggesting that the odds of a borrower connected
to exactly one loan purpose is approximately `26.34` times higher than
the odds of observing a configuration where the degree is not one. In
other words, in this Bondora sample, it is highly likely that borrowers
select only one loan type. This effect size or error has not deviated
strongly across the four model permutations, indicating its stability.
This result shows support for the Theory of Perceived Diversification by
@ayal_perceived_2009 because on an **individual level,** borrowers
prefer to hold minimal *distinct* debt types.

Further, the third hypothesis asserts that borrowers affiliated by a
common loan purpose overlap on a second loan purpose, reflecting
familiarity bias on an **aggregated** level. In Model 4, four-cycles are
insignificant, meaning that we have not found sufficient evidence to
reject the null hypothesis that there are no systematic co-affiliation
effects. In other words, borrowers who are affiliated by the same loan
purpose do not overlap on a second loan purpose. This finding is
contrary to @herzenstein_strategic_2011, who argues that borrowers
aggregate around popular, or familiar loans. This could be explained by
the fact that the authors examine auction bids, rather than winning
bids, as we have. The behaviours or characteristics of bidders, compared
to winners, may be fundamentally different.

Evaluating attribute-level effects, we see that neither age nor gender
has a significant impact on loan purpose selection by borrowers.
Specifically, we do not find sufficient evidence to conclude that
borrowers being younger has any influence over how they select their
loan purposes. This indicates a lack of support for hypothesis 5, which
asserts that younger borrowers should have different loan purpose
combinations than older borrowers. In this vein, we cannot corroborate
@cooper_consumption_2023, however, this is likely due to the fact that
their study focuses on credit consumption of U.S. young adults, which is
fundamentally different than credit consumption in Europe
@jentzsch_consumer_2006. Moreover, they studied traditional credit
markets such as personal credit lines from banks rather than
peer-to-peer lending services, which also vary fundamentally.

In the same vein, we do not observe sufficient evidence to determine
that borrowers' gender affects how they select their specific loan
purposes. Similarly, this is contrary to hypothesis 6, which states that
there is a gender difference in how borrowers select loan purposes. The
lack of effect is contrary to @aliano_role_2023, who finds gender to be
a significant predictor of loan purpose heterogeneity. They argue that
gender differences across credit risk perception and self-selection
effects are responsible, however, it is likely that these behavioural
factors are already captured by the `b1degree(1)` term; as
@ayal_perceived_2009 argued, similar risk-related phenomena are
responsible for Prospect Theory, which in turn drives the Theory of
Perceived Diversification.

```{r fig.width=10, fig.height=5}
#| label: fig-ergm-coef
#| fig-cap: "ERGM Coefficient Plots"

ep1 <- readRDS(here::here("resources","objects","ergm","m1_coef_plot.Rds"))
ep2 <- readRDS(here::here("resources","objects","ergm","m2_coef_plot.Rds"))
ep3 <- readRDS(here::here("resources","objects","ergm","m3_coef_plot.Rds"))
ep4 <- readRDS(here::here("resources","objects","ergm","m4_coef_plot.Rds"))

gridExtra::grid.arrange(ep1, ep2, ep3, ep4, nrow = 2, ncol = 2)
```

```{r}
#| output: asis
ergm_table_items <- readRDS(here::here("resources","objects","ergm",                                 "ergm_table.Rds"))

texreg::texreg(ergm_table_items[[1]], 
               custom.model.names = ergm_table_items[[2]],
               digits = 3,
               caption.above = TRUE,
               caption = "ERGM Modelling Results",
               label = "tab-ergm-results",
               booktabs = TRUE,
               float.pos = "!h",
               use.packages = FALSE,
               scalebox = 0.85,
               dcolumn = TRUE)
```

Evaluating model fit, we must first evaluate the MCMC simulations seen
in Figure below. Observing Model 4, we see that all terms have mixed
sufficiently well. While `edges`, `b1degree(1)`, `b1factor("gender")`,
and `b1cov("age")` are well centred at zero, `cycle(4)` is slightly
right-skewed, indicating minor mixing issues. Additionally, the trace
plots appear to resemble a "fuzzy" shape, supporting this. We find this
diagnostic acceptable.

Further, the GOF measures indicate how well simulation iterations match
our observed network. Across the GOF measures, p-values close to `1`
resemble a close match with reality. All geodesic distances except for
`Inf` and `2` are well-modelled. The former is expected, as our
bipartite configuration restricts any possibility of links within each
partition, meaning that `Inf` distances can never be modelled. For
`b1degree(1)`, all degrees except for zero and two are highly
insignificant. It is expected that zero degrees is poorly modelled
because our observed network has no isolates. Degree three is arguably
poorly modelled since there are only `7` observations of it, limiting
the simulations. The term `b2degree()` is not modelled well, however,
this term is very difficult to capture statistically; there is no
meaningful statistical variation with respect how many borrowers each
loan purpose has, so the GOF test cannot reliably evaluate it. Moreover,
edge-wise shared partners are all well modelled. Overall, we consider
the GOF tests to be sufficient.

Finally, we evaluate the optimal statistical model configuration by
observing the minimum AIC/BIC values across the models. AIC prefers
Model 3 whereas BIC chooses Model 1; this is expected, as BIC penalizes
less informative model terms more strongly.

```{r fig.width=10, fig.height=5, echo=FALSE,results='hide',fig.keep='all'}
#| label: fig-m4-diags
#| fig-cap: "Model 4 Diagnostics Plots"

em4 <- readRDS(here::here("resources","objects","ergm","ergm_m4_panel.Rds"))

invisible(ergm::mcmc.diagnostics(em4$model))
invisible(snafun::stat_plot_gof(em4$gof))
```

## Conclusion

This study aimed to assess the prevalence of behavioural biases in a
peer-to-peer lending context by assessing how connections across
borrowers and loan types varied; by answering the question *"Do
borrowers vary in their loan purpose type and frequency based on their
individual and loan characteristics"*, we can determine the extent that
borrowers have an aversion to large loan counts within loan purpose
categories, per @ayal_two_2011, and by answering *"What behavioural
mechanisms affect the structure of borrower-to-loan-use connections?"*
in a network setting, we can see how loan purpose choices vary with
structural factors.

To evaluate how loan counts are affected within loan purpose selections,
we utilised the QAP linear regression framework. Within this study,
borrowers are studied as dyads rather than a whole network. Firstly, it
was found that borrowers having the same credit rating was significant
in explaining variations in the number of loans they took out for a
shared purpose, however, the impact was small in size. Further, we found
that a shared occupation area between borrowers does not have a
meaningful impact on loan counts. Assessing the models, we find a poor
linear fit, which could indicate that dyads are not an optimal means for
studying these effects. These findings do not necessarily corroborate
@ayal_two_2011 because we do not observe strong negative effects on loan
counts, however, the regression diagnostics make this conclusion quite
unreliable.

To evaluate specific borrower behaviours independent of loan counts, a
network statistical model such as ERGM is more suitable. In this study,
we differentiate between **aggregate** and **individual** borrower level
effects. The results indicate very strongly that borrowers prefer to
hold one loan type. This yields significant support to the Theory of
Perceived Diversification by @ayal_perceived_2009, capturing the idea
that on an **individual** level, borrowers try to have the least amount
of distinctiveness possible. Specifically, by holding only one loan
type, they can place them in the same mental 'account', which prevents
distress [@ayal_perceived_2009]. However, we do not find that, in
**aggregate,** borrowers have co-affiliation by being related to two
distinct loan types. In other words, borrowers do not overlap with their
choices of pairs of loan types. These two findings highlight that
borrowers are more prone to individual behavioural biases rather than a
'herding' effect as described by @herzenstein_strategic_2011. However,
as we do not utilise auction data, it is possible that aggregate effect
does not extend to how *winners* of auctions behave.

Further, examining borrowers' attribute effects, we found that neither
gender nor age have significant effects on how borrowers choose loan
types. Specifically, gender differences do not account for a difference
in loan purpose formation. We argue that this could be due to the idea
that the same gender differences that affect risk perception are also
responsible for the underlying behavioural phenomena behind the Theory
of Perceived Diversification. Consequently, the use of `b1degree(1)` had
already captured this process. Finally, we do not find that younger
borrowers have a different loan purpose mix than older borrowers. This
theory was derived from @cooper_consumption_2023, who studied young
adults in the U.S. However, considering the fundamental differences
across credit markets between the U.S. and Europe, it is likely that
these behavioural effects may not apply to European borrowers across our
sample of three countries. Evaluating the ERGM models, we find them to
be well constructed and our diagnostics show that they largely conform
to our expectations. Therefore, we consider these findings to be robust
to modelling or specification error.

Arguably, the main beneficiaries of our findings are credit regulatory
agencies in Finland, Estonia, and Spain, to which the majority of our
borrowers fall under. Some studies indicate shortcomings in how these
specific credit markets are regulated, owing their young age and online
nature [@lenzpeerpeer2016]. By understanding the behavioural phenomena
driving borrowers' behaviour, these institutions can construct policies
to minimise the risks of users utilising credit inappropriately, for
example, by systematically maximizing credit counts. Further, while we
have studied how credit counts and borrower selections are influenced,
we have been unable to combine these phenomena together by investigating
loan counts within the ERGM framework. This extension would be valuable
because it would account for the shortcomings of the QAP regressions'
linearity whilst also allowing for more complex behavioural phenomena to
be modelled. In general, however, we have observed that researchers
focus on singular P2P platforms. We recommend that studies focus on
concurrent analyses of a combination of P2P platforms because individual
platforms may have idiosyncrasies that make analyses difficult to
generalise to the wider P2P lending market.

\clearpage

## References {.unnumbered}

::: {#refs}
:::

\clearpage
\appendix

## Appendix

\label{sec:supp}
 \definecolor{dkgreen}{rgb}{0,0.6,0} \definecolor{gray}{rgb}{0.5,0.5,0.5} \definecolor{mauve}{rgb}{0.58,0,0.82} \lstset{ %
  language=R,                     % the language of the code
  basicstyle=\footnotesize,       % the size of the fonts that are used for the code
  numbers=left,                   % where to put the line-numbers
  numberstyle=\tiny\color{gray},  % the style that is used for the line-numbers
  stepnumber=1,                   % the step between two line-numbers. If it's 1, each line
                                  % will be numbered
  numbersep=5pt,                  % how far the line-numbers are from the code
  backgroundcolor=\color{white},  % choose the background color. You must add 
  showspaces=false,               % show spaces adding particular underscores
  showstringspaces=false,         % underline spaces within strings
  showtabs=false,                 % show tabs within strings adding particular underscores
  frame=false,                   % adds a frame around the code
  rulecolor=\color{black},        % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. commens (green here))
  tabsize=2,                      % sets default tabsize to 2 spaces
  captionpos=b,                   % sets the caption-position to bottom
  breaklines=true,                % sets automatic line breaking
  breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
  title=\lstname,                 % show the filename of files included with \lstinputlisting;
                                  % also try caption instead of title
  keywordstyle=\color{blue},      % keyword style
  commentstyle=\color{dkgreen},   % comment style
  stringstyle=\color{mauve},      % string literal style
}

### Data Preprocessing Steps

\label{sec:preprocessing}

This section outlines the specific steps taken to process the data into
the network object that was used in later analysis, outlining the
rationale, where needed.

-   Following the data import, only the following attributes were kept
    to make our processing steps more focused

``` r
keep_cols <- c("LoanId", "UserName","Age", "Gender", 
               "Country", "Amount", "Interest","LoanDuration",
              "UseOfLoan", "Rating", "Restructured", "MonthlyPayment",
              "OccupationArea", "BiddingStartedOn")
```

-   We removed rows with any `NA` values for the selected attributes to
    ensure that we have a complete dataset

-   Observing @fig-user-counts, we see that between 2017-2021, there
    exists many `-1` values for `UseofLoan`. This is because Bondora no
    longer provided this detail in later datasets. However, between 2014
    and 2016 this data is available. Because of this, we restrict the
    time period to have the least noisy data. `BiddingStartedOn` was
    chosen to record the time, because the auction start date can be a
    good indication of activity around the loan.

```{r fig.width=14, fig.height=5}
#| label: fig-user-counts
#| fig-cap: "Number of Unique Users per Loan Type per Year"

replayPlot(readRDS(here::here("resourceS","objects","preprocessing","user_cnt_plt.Rds")))
```

-   Then, we sampled `500` individuals out of the filtered dataset to
    ensure that our models run sufficiently well and converge. By
    choosing `500` individuals, we preserve importance structures within
    the network while adding sufficient statistical power.

<!-- -->

-   When adding labels to each loan purpose and occupation area, we keep
    any possible missing values because these do not indicate lower
    quality data. Instead, they can carry information about lenders
    distribute loans according to the quality of information presented
    by borrowers [@yao_relationship_2019]. Additionally, keeping missing
    labels is important to ensure that the edges are not superficially
    limited and that nodes can be isolates if they are indeed that way
    in reality.

-   To create the network object for ERGM models:

    -   We first create an incidence matrix of size $n \times m$,
        borrower usernames by the loan purpose, where the values are the
        number of loans the user obtained for each loan purpose
        category. These values are then turned into binary because our
        modelling does not support the use of weighted edges.

    -   A bipartite network is created from this incidence matrix, where
        the first partition are unique users, and the second partition
        is the loan purpose associated among all loans of the user.
        Users can share more than more loan purpose type.

    -   We add age and gender attributes to each node of the first
        partition

-   To create the relevant networks/matrices for QAP Regressions:

    -   We first re-create the loan purpose incidence matrix, however,
        we keep the counts as the linear regression supports this.

        -   We then convert the loan purpose incidence matrix into an
            adjacency matrix of shape $n\times n$ through the
            transformation $\mathbf{X}\cdot\mathbf{X^T}$, which results
            in a matrix of shared loan purposes weighted by the
            frequency of each user's loan count within each loan purpose
            category.

        -   We then set the diagonal values of $\mathbf{X}$ to be zero
            to ensure that there are no loops within the network.

    -   In accordance with this procedure, we create adjacency matrices
        for `Credit Rating` and `Occupation Area` which will be used as
        the main predictors in the QAP models.

        -   `Credit Rating` is a weighted adjacency matrix based on the
            loan count belonging to each user for each credit rating
            category

        -   `Occupation Area` is a binary adjacency matrix based on
            which occupation each user has reported to have held in the
            past when applying for each of their loans.

    -   Additionally, we create a set of control variables to improve
        the validity and performance of the linear regression using
        `Loan Amounts`, `Age`, `Gender`, `Loan Duration`, and whether
        the loans have been restructured.

        -   `Loan Amounts` is a weighted adjacency matrix based on five
            bins across the range of possible loan amounts

        -   `Age` is an adjacency matrix based on differences in users'
            ages

        -   `Gender` is a binary adjacency matrix based on users' shared
            gender.

        -   `Loan Duration` is an adjacency matrix based on the
            differences in users' loans average durations

        -   `Restructured` is a binary adjacency matrix based on whether
            the users share the fact that they have defaulted on any
            loans in the past. The users can either have shared both
            defaults and non-defaults, or both.

\clearpage

### Distributions of Variables across Samples

\label{sec:dist}

```{r fig.width=14, fig.height=5}
#| label: fig-hist-amt
#| fig-cap: "Distribution of Loan Amount across Samples"
#| warning: false

replayPlot(readRDS(here::here("resources","objects","preprocessing",
                          "hist_amt.Rds")))
```

```{r fig.width=14, fig.height=5}
#| label: fig-hist-int
#| fig-cap: "Distribution of Interest Rates across Samples"
#| warning: false

replayPlot(readRDS(here::here("resources","objects","preprocessing",
                          "hist_int.Rds")))
```

```{r fig.width=14, fig.height=5}
#| label: fig-hist-age
#| fig-cap: "Distribution of Age across Samples"
#| warning: false

replayPlot(readRDS(here::here("resources","objects","preprocessing",
                          "hist_age.Rds")))
```

```{r fig.width=14, fig.height=5}
#| label: fig-hist-loandur
#| fig-cap: "Distribution of Loan Duration across Samples"
#| warning: false

replayPlot(readRDS(here::here("resources","objects","preprocessing",
                          "hist_loandur.Rds")))
```

```{r fig.width=14, fig.height=5}
#| label: fig-bar-loanuse
#| fig-cap: "Distribution of Loan Purpose across Samples"
#| warning: false

replayPlot(readRDS(here::here("resources","objects","preprocessing",
                          "bar_loanuse.Rds")))
```

```{r fig.width=14, fig.height=5}
#| label: fig-bar-rating
#| fig-cap: "Distribution of Credit Ratings across Samples"
#| warning: false

replayPlot(readRDS(here::here("resources","objects","preprocessing",
                          "bar_rating.Rds")))
```

```{r fig.width=14, fig.height=7}
#| label: fig-bar-occupation
#| fig-cap: "Distribution of Occupation Types across Samples"
#| warning: false

replayPlot(readRDS(here::here("resources","objects","preprocessing",
                          "bar_occupation.Rds")))
```

```{r fig.width=14, fig.height=7}
#| label: fig-bar-country
#| fig-cap: "Distribution of Borrower Origins"
#| warning: false

replayPlot(readRDS(here::here("resources","objects","preprocessing",
                          "bar_country.Rds")))
```

\clearpage

### ERGM Results - Supplementary Tables

```{r}
#| label: tbl-ergm-prob-tables
#| tbl-cap: "ERGM Results in Probability Scale"
knitr::kable(readRDS(
  here::here("resources","objects","ergm","ergm_prob_tables.Rds")))
```

\clearpage

### ERGM Diagnostics Plots

\label{sec:ergm_diags}

#### Base Model

\begingroup
\scriptsize

```{r fig.width=10, fig.height=5}

base <- readRDS(here::here("resources","objects","ergm","ergm_base_panel.Rds"))

snafun::stat_plot_gof(base$gof)
base$gof
```

#### Model 1

```{r fig.width=10, fig.height=5}

em1 <- readRDS(here::here("resources","objects","ergm","ergm_m1_panel.Rds"))

ergm::mcmc.diagnostics(em1$model)
snafun::stat_plot_gof(em1$gof)
em1$gof
```

#### Model 2

```{r fig.width=10, fig.height=5}

em2 <- readRDS(here::here("resources","objects","ergm","ergm_m2_panel.Rds"))

ergm::mcmc.diagnostics(em2$model)
snafun::stat_plot_gof(em2$gof)
em2$gof
```

#### Model 3

```{r fig.width=10, fig.height=10}

em3 <- readRDS(here::here("resources","objects","ergm","ergm_m3_panel.Rds"))

ergm::mcmc.diagnostics(em3$model)
```

```{r fig.width=10, fig.height=5}

snafun::stat_plot_gof(em3$gof)
em3$gof
```

#### Model 4

This section contains only the console output for the GOF diagnostics
for Model 4. The main diagnostics are contained in the Results section.

```{r fig.width=10, fig.height=10}

em4$gof
```

\clearpage
\endgroup

### Source Code - Data Preprocessing

\linespread{1} 
\lstinputlisting[language=R]{scripts/bondora_preprocessing.R}
\clearpage

### Source Code - QAP Linear Regression

\linespread{1} 
\lstinputlisting[language=R]{scripts/qap_network_analysis.R} 
\clearpage

### Source Code - ERGM Network Analysis

\linespread{1} 
\lstinputlisting[language=R]{scripts/ergm_network_analysis.R} 
\clearpage
\linespread{1.15}

## Technology Statement

> During the preparation of this work, we used ChatGPT in order to
> generate select parts of the `R` script utilised to process the
> dataset. Specifically, the tool was used to transform the processed
> dataset into a format that `igraph` would accept as a network object.
> Additionally, some `R` functions written by the group were improved
> with ChatGPT's suggestions. No AI tool was utilised to independently
> write parts of the report. The following parts of the assignment were
> affected/generated by AI tool usage: **INTRODUCTION;** The tool was
> utilised to evaluate the validity of certain theoretical concepts and
> refine them. **DATASET;** the data described within this section was
> processed partly by some code drafted by ChatGPT and edited by the
> group. After using this tool/service, **Group 7** evaluated the
> validity of the tool’s outputs, including the sources that generative
> AI tools have used, and edited the content as needed. As a
> consequence, **Group 7** takes full responsibility for the content of
> their work.
