---
title: "Bondora Credit Allocation Study"
subtitle: "Progress Meeting 2"
author: "Group 7"
format:
  revealjs: 
    embed-resources: true
    self-contained-math: true 
    center-title-slide: true
    self-contained: true
    theme: dark
    scrollable: true
---

```{r include=FALSE}
#install.packages("here") # if you do not have this installed already

net <- readRDS(here::here("resources","objects","preprocessing","bondora_net.RDS"))
b_indicator <- readRDS(here::here("resources","objects","preprocessing","indicator.RDS"))

base_ergm <- readRDS(here::here("resources","objects","ergm","ergm_base_panel.RDS"))
ergm_m1 <- readRDS(here::here("resources","objects","ergm","ergm_m1_panel.RDS"))
ergm_m2 <- readRDS(here::here("resources","objects","ergm","ergm_m2_panel.RDS"))
ergm_m3 <- readRDS(here::here("resources","objects","ergm","ergm_m3_panel.RDS"))
ergm_m4 <- readRDS(here::here("resources","objects","ergm","ergm_m4_panel.RDS"))

qap_m1 <- readRDS(here::here("resources","objects","qap","qap_m1.RDS"))
qap_m2 <- readRDS(here::here("resources","objects","qap","qap_m2.RDS"))
```

# How are Different Kinds of Credit Distributed among Bondora Borrowers?

## Background & Premise {.smaller}

-   Bondora P2P Lending Service
-   European dataset spanning between 2009-2021 with 179,000+ individual loans
-   Each user can have 1+ loan

### Network Information

+------------------+------------------------------------------------------------------+
| **Network Type** | **Undirected Bipartite**                                         |
|                  |                                                                  |
| Partition 1 (B1) | Borrowers (based on usernames)                                   |
|                  |                                                                  |
| Partition 2 (B2) | Reported Loan Use                                                |
+------------------+------------------------------------------------------------------+
| **Edges**        | B1 $\rightarrow$ B2 if user took out loan for a specific purpose |
+------------------+------------------------------------------------------------------+

## Research Questions & Hypotheses {.smaller}

::: panel-tabset
### Study 1

+------------+-------------------------------------------------------------------------------------+
| **Method** | QAP Linear Regression                                                               |
+------------+-------------------------------------------------------------------------------------+
| **RQ**     | How do loan characteristics affect the frequency credit allocation among borrowers? |
+------------+-------------------------------------------------------------------------------------+

+----------------------------------------------------------------------------------+--------------------------------------------------+-------------------------------------------------------------+
| Hypothesis                                                                       | Specific Method                                  | Literature / Theory                                         |
+==================================================================================+==================================================+=============================================================+
| H~1~^a^ : Users with the same loan use are likely to have the same credit rating | $Y$ : Adjacency matrix of shared loan use        | Lorem Ipsum Lorem Ipsum Lorem Ipsum Lorem Ipsum Lorem Ipsum |
|                                                                                  |                                                  |                                                             |
| We expect $\hat{\beta}_{1,1}>0$                                                  | $X_1$ : Adjacency matrix of shared credit rating |                                                             |
|                                                                                  |                                                  |                                                             |
|                                                                                  | $+$ $X_{2,3,..,n}$ : Controls                    |                                                             |
+----------------------------------------------------------------------------------+--------------------------------------------------+-------------------------------------------------------------+

### Study 2

+------------+-----------------------------------------------------------------------------------------------------+
| **Method** | Bipartite ERGMs                                                                                     |
+------------+-----------------------------------------------------------------------------------------------------+
| **RQ**     | How is the structure credit type allocation affected by borrower behaviour and its characteristics? |
+------------+-----------------------------------------------------------------------------------------------------+

+--------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------+
| Hypothesis                                                                     | ERGM Term                                                                                                             | Literature / Theory                                         |
+================================================================================+=======================================================================================================================+=============================================================+
| H~2~^a^ : Pairs of borrowers share same loan type                              | Endogenous + Dyad-dependent                                                                                           | Lorem Ipsum Lorem Ipsum Lorem Ipsum Lorem Ipsum Lorem Ipsum |
|                                                                                |                                                                                                                       |                                                             |
| We expect $\hat{\theta}>0$                                                     | `gwb1dsp()`                                                                                                           |                                                             |
+--------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------+
| H~3~^a^ : Borrowers have a diverse loan use portfolio                          | Endogenous + Dyad-dependent                                                                                           | Lorem Ipsum Lorem Ipsum Lorem Ipsum Lorem Ipsum Lorem Ipsum |
|                                                                                |                                                                                                                       |                                                             |
| We expect $\hat{\theta}>0$                                                     | `gwb1degree()`                                                                                                        |                                                             |
+--------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------+
| H~4~^a^ : Younger borrowers have a more diverse range of loan uses             | Exogenous + Dyad-independent                                                                                          | Lorem Ipsum Lorem Ipsum Lorem Ipsum Lorem Ipsum Lorem Ipsum |
|                                                                                |                                                                                                                       |                                                             |
| We expect $\hat{\theta}>0$                                                     | `b1cov("b1_age")`                                                                                                     |                                                             |
+--------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------+
| H~5~^a^ : Borrowers of differing genders have different portfolios of loan use | Heterophily + Exogenous + Dyad-independent                                                                            | Lorem Ipsum Lorem Ipsum Lorem Ipsum Lorem Ipsum Lorem Ipsum |
|                                                                                |                                                                                                                       |                                                             |
| We expect $\hat{\theta}<0$                                                     | `b1nodematch("b1_gender")`                                                                                            |                                                             |
|                                                                                |                                                                                                                       |                                                             |
|                                                                                | **Note:** measured indirectly through heterophily because `nodemix()` does not have an equivalent for bipartite ERGMs |                                                             |
+--------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------+
:::

## Basic Network Statistics {.smaller}

::: panel-tabset
### Network Visualisation

```{r}
# Set colour palette
cols <- viridis::viridis(30)

# Copy network for plotting
bondora_plot <- net

# Get node type for plotting
type_indicator <- ifelse(b_indicator == 2, TRUE,FALSE)
shape <- ifelse(type_indicator,"square","circle")
network::set.vertex.attribute(bondora_plot, "shape", shape)

# Get Category Count for Vertex Size
counts <- summary(bondora_plot ~ b2sociality)
counts_att <- ifelse(type_indicator, counts^0.75, counts^0.6)
network::set.vertex.attribute(bondora_plot, "size", counts_att)

# Colours for the Node Types
plot_cols <- ifelse(type_indicator, cols[5], cols[30])
network::set.vertex.attribute(bondora_plot, "color", plot_cols)

# Legend Plotting
type_legend <- ifelse(type_indicator, "Borrowers", "Loan Type")
type_legend <- as.factor(type_legend)

# Plot the Network
plot(snafun::to_igraph(bondora_plot),
     main = "Bipartite User-LoanUse",
     edge.arrow.size = 0.3,
     edge.color = rgb(0,0,0, alpha = 0.35),
     vertex.frame.color = "black",
     vertex.label = NA,
     vertex.frame.size = 3,
     edge.curved = FALSE,
     layout=igraph::layout.fruchterman.reingold)
legend("bottomleft", legend = levels(type_legend), 
       inset = c(0.1, 0.02),
       col = c(cols[5], cols[30]),
       pch = c(16,15), 
       title = "Node Partitions", title.font = 2,
       cex = 0.8,       
       lwd = 1,
       bg = rgb(0,0,0, alpha=0.025))
```

### Descriptive Statistics Plot

```{r}
snafun::plot_centralities(net)
```
:::

## Study 1 \| Preliminary QAP Results {.smaller}

```{r include=FALSE}
var_names <- c("Intercept", "Rating","Loan Amount","Age","Gender",
               "Loan Duration","Restructured")
qap1_results <- qap_m1$coefficients
qap2_results <- qap_m2$coefficients
names(qap1_results) <- var_names
names(qap2_results) <- var_names
```

::: panel-tabset
### Notes

-   Linear QAP Regression used to study how borrowers with the same loan type choice are affected by having the same credit rating

-   The adjacency matrices used are **weighted** to give this study a level of detail we do not have with ERGM

    -   Because the weights vary significantly, **standardised** results are more meaningful to compare

-   Intercept has very significant impact: *not all effects can be accounted for and are missing*

### Parameters

+--------------------------+------------------------------------------------------------------+
| Variable Type            | Variable Name                                                    |
+==========================+==================================================================+
| **Dependent Variable**   | Weighted Adjacency Matrix of Borrowers with same Loan Type       |
+--------------------------+------------------------------------------------------------------+
| **Explanatory Variable** | Weighted Adjacency Matrix of Borrowers of the same Credit Rating |
+--------------------------+------------------------------------------------------------------+
| Control Variable         | Weighted Adjacency Matrix of Mean Loan Amount Differences        |
+--------------------------+------------------------------------------------------------------+
| Control Variable         | Weighted Adjacency Matrix of Age Differences                     |
+--------------------------+------------------------------------------------------------------+
| Control Variable         | Binary Adjacency Matrix of Homophily in Gender                   |
+--------------------------+------------------------------------------------------------------+
| Control Variable         | Weighted Adjacency Matrix of Mean Loan Duration Differences      |
+--------------------------+------------------------------------------------------------------+
| Control Variable         | Weighted Adjacency Matrix of Homophily in Loan Restructuring     |
+--------------------------+------------------------------------------------------------------+

### Specification & Results

``` r
qap_m1 <- sna::netlm(y = loan_use_mat,
                     x = list(rating_mat, amt_diffs_mat, age_diffs_mat,
                              gender_mat, loandur_diffs_mat, rest_mat),
                     nullhyp = "qapspp", reps = 2500)

# Standardised QAP Linear Regression
scaled_dep <- scale(loan_use_mat)
scaled_pred <- lapply(pred_vars, scale)

qap_m2 <- sna::netlm(y = scaled_dep,
                     x = scaled_pred,
                     nullhyp = "qapspp", reps = 2500)
```

```{r}
coef_table <- cbind(Unstd_Parameters = round(qap1_results, 4),
                    Ustd_t_vals = round(qap_m1$tstat, 2),
                    Std_Parameters = round(qap2_results, 4),
                    Std_t_vals = round(qap_m2$tstat, 2))
print(coef_table)
```
:::

## Study 2 \| ERGM \| `edges` {.smaller}

::: panel-tabset
### Notes

-   Sparse edges but bipartite, so take `edges` with grain of salt

-   Unconstrained

-   GOF acceptable for an initial non-tuned ERGM

### Specification & Summary

``` r
formula_base_model <- bondora_net ~ edges
base_ergm <- ergm::ergm(formula_base_model)
```

```{r}
texreg::screenreg(list(base_ergm$model))
```

### GOF

```{r}
snafun::stat_plot_gof(base_ergm$gof)
```
:::

## Study 2 \| ERGM \| `gwb1degree` {.smaller}

::: panel-tabset
### Notes

-   `gwb1degree` refers to number of different types of loans users request

### Specification

``` r
model_1_params <- bondora_net ~ edges + 
  # b1 decay can be very low since 9 b2
  gwb1degree(decay=0.15, fixed=TRUE) 

model_1 <- ergm::ergm(
  model_1_params,
  
  # Max b2 degree is 72, so this constraint is reasonable
  # and helps convergence significantly.
  # Technically in the Bondora population this can be 
  # far higher but we are studying a subsample.
  #constraints = ~ bd(minout = 0, maxout = 80),
  
  control = ergm::control.ergm(
    # Greater burn-in for cleaner result
    MCMC.burnin = 20000,
    # Greater sample size for greater stability
    MCMC.samplesize = 100000,
    seed = 42,
    MCMC.interval = 1000,
    # Only needed for convergence pvals to improve
    MCMLE.maxit = 45,
    # Smaller steps for stability
    MCMLE.steplength = 0.25,
    parallel = n_cores,
    parallel.type = "PSOCK"
  )
)
```

### Results

```{r}
texreg::screenreg(list(
  base_ergm$model, ergm_m1$model))
```

### MCMC

```{r}
ergm::mcmc.diagnostics(ergm_m1$model)
```

### GOF

```{r}
snafun::stat_plot_gof(ergm_m1$gof)
```
:::

## Study 2 \| ERGM \| `gwb1degree + gwb1dsp` {.smaller}

::: panel-tabset
### Notes

-   Constrained model

### Specification

``` r
model_2_params <- bondora_net ~ edges + 
  # low decay important because there is high clustering around low degrees
  gwb1degree(decay=0.15, fixed=TRUE) + 
  # decay should be higher due to wider variation in degree but 
  # too high of degree makes the traces concentrated around the tails.
  gwb1dsp(decay=0.5, fixed=TRUE)      
  
model_2 <- ergm::ergm(
  model_2_params,
  
  # Max b2 degree is 72, so this constraint is reasonable
  # and helps convergence significantly.
  # Technically in the Bondora population this can be 
  # far higher but we are studying a subsample.
  constraints = ~ bd(minout = 0, maxout = 80),
  
  control = ergm::control.ergm(
    # Greater burn-in for cleaner result
    MCMC.burnin = 20000,
    # Greater sample size for greater stability
    MCMC.samplesize = 100000,
    seed = 42,
    MCMC.interval = 1000,
    # Only needed for convergence pvals to improve
    MCMLE.maxit = 45,
    # Smaller steps for stability
    MCMLE.steplength = 0.25,
    parallel = n_cores,
    parallel.type = "PSOCK"
    )
  )
```

### Results

```{r}
texreg::screenreg(list(base_ergm$model, ergm_m1$model, 
                       ergm_m2$model))
```

### MCMC

```{r}
ergm::mcmc.diagnostics(ergm_m2$model)
```

### GOF

```{r}
snafun::stat_plot_gof(ergm_m2$gof)
```
:::

## Study 2 \| ERGM \| `gwb1degree + gwb1dsp + b1nodematch("gender")` {.smaller}

::: panel-tabset
### Notes

-   Constrained model

### Specification

``` r
model_3_params <- bondora_net ~ edges + 
  # low decay important because there is high clustering around low degrees
  gwb1degree(decay=0.15, fixed=TRUE) + 
  # decay should be higher due to wider variation in degree but 
  # too high of degree makes the traces concentrated around the tails.
  gwb1dsp(decay=0.5, fixed=TRUE) +
  # See differences across genders (implicitly, since b1nodemix unavailable)
  b1nodematch("b1_gender", diff=FALSE)

model_3 <- ergm::ergm(
  model_3_params,
  
  # Max b2 degree is 72, so this constraint is reasonable
  # and helps convergence significantly.
  # Technically in the Bondora population this can be 
  # far higher but we are studying a subsample.
  constraints = ~ bd(minout = 0, maxout = 80),
  
  control = ergm::control.ergm(
    # Greater burn-in for cleaner result
    MCMC.burnin = 20000,
    # Greater sample size for greater stability
    MCMC.samplesize = 100000,
    seed = 42,
    MCMC.interval = 1000,
    # Only needed for convergence pvals to improve
    MCMLE.maxit = 45,
    # Smaller steps for stability
    MCMLE.steplength = 0.25,
    parallel = n_cores,
    parallel.type = "PSOCK"
  )
)
```

### Results

```{r}
texreg::screenreg(list(base_ergm$model, ergm_m1$model, 
                       ergm_m2$model,ergm_m3$model))
```

### MCMC

```{r}
ergm::mcmc.diagnostics(ergm_m3$model)
```

### GOF

```{r}
snafun::stat_plot_gof(ergm_m3$gof)
```
:::

## Study 2 \| ERGM \| `gwb1degree + gwb1dsp + b1nodematch("gender") + b1cov("age")` {.smaller}

::: panel-tabset
### Notes

-   Constrained model

### Specification

``` r
model_4_params <- bondora_net ~ edges + 
  # low decay important because there is high clustering around low degrees
  gwb1degree(decay=0.15, fixed=TRUE) + 
  # decay should be higher due to wider variation in degree but 
  # too high of degree makes the traces concentrated around the tails.
  gwb1dsp(decay=0.5, fixed=TRUE) +
  # See differences across genders (implicitly, since b1nodemix unavailable)
  b1nodematch("b1_gender", diff=TRUE) +
  # See if higher ages make a difference
  b1cov("b1_age")

model_4 <- ergm::ergm(
  model_4_params,
  
  # Max b2 degree is 72, so this constraint is reasonable
  # and helps convergence significantly.
  # Technically in the Bondora population this can be 
  # far higher but we are studying a subsample.
  constraints = ~ bd(minout = 0, maxout = 80),
  
  control = ergm::control.ergm(
    # Greater burn-in for cleaner result
    MCMC.burnin = 20000,
    # Greater sample size for greater stability
    MCMC.samplesize = 100000,
    seed = 42,
    MCMC.interval = 1000,
    # Only needed for convergence pvals to improve
    MCMLE.maxit = 45,
    # Smaller steps for stability
    MCMLE.steplength = 0.25,
    parallel = n_cores,
    parallel.type = "PSOCK"
  )
)
```

### Results

```{r}
texreg::screenreg(list(base_ergm$model, ergm_m1$model, 
                       ergm_m2$model, ergm_m3$model, 
                       ergm_m4$model))
```

### MCMC

```{r}
ergm::mcmc.diagnostics(ergm_m4$model)
```

### GOF

```{r}
snafun::stat_plot_gof(ergm_m4$gof)
```
:::
