---
title: "Bondora Credit Allocation Study"
subtitle: "Progress Meeting 2"
author: "Group 7"
bibliography: references.bib
csl: apa-6th-edition.csl
institute: "Jheronimus Academy of Data Science"
execute:
  cache: true
format:
  revealjs: 
    margin: 0.05
    embed-resources: true
    self-contained-math: true 
    center-title-slide: true
    self-contained: true
    theme: simple 
    scrollable: true
---

```{r include=FALSE}
#install.packages("here") # if you do not have this installed already

net <- readRDS(here::here("resources","objects","preprocessing","bondora_net.Rds"))

base_ergm <- readRDS(here::here("resources","objects","ergm","ergm_base_panel.Rds"))
ergm_m1 <- readRDS(here::here("resources","objects","ergm","ergm_m1_panel.Rds"))
ergm_m2 <- readRDS(here::here("resources","objects","ergm","ergm_m2_panel.Rds"))
ergm_m3 <- readRDS(here::here("resources","objects","ergm","ergm_m3_panel.Rds"))
ergm_m4 <- readRDS(here::here("resources","objects","ergm","ergm_m4_panel.Rds"))

qap_m1 <- readRDS(here::here("resources","objects","qap","qap_m1.Rds"))
qap_m2 <- readRDS(here::here("resources","objects","qap","qap_m2.Rds"))

h1 <- readRDS(here::here("resources","objects","preprocessing","hist_amt.Rds"))
h2 <- readRDS(here::here("resources","objects","preprocessing","hist_int.Rds"))
h3 <- readRDS(here::here("resources","objects","preprocessing","hist_loandur.Rds"))
h4 <- readRDS(here::here("resources","objects","preprocessing","hist_monpmt.Rds"))
h5 <- readRDS(here::here("resources","objects","preprocessing","hist_age.Rds"))

br1 <- readRDS(here::here("resources","objects","preprocessing","bar_loanuse.Rds"))
br2 <- readRDS(here::here("resources","objects","preprocessing","bar_rating.Rds"))

qap_plt <- readRDS(here::here("resources","objects","qap","unstd_std_plot.Rds"))
ergm_plt <- readRDS(here::here("resources","objects","ergm","network_plot.Rds"))
```

```{r include=FALSE}
prob <- function(input) {
  return(unname(round(exp(input)/(1+exp(input))),2))
}
odds <- function(input) {
  return(unname(round(exp(input)/(1+exp(input)),2)))
}
```

# How are Different Kinds of Credit Distributed among Bondora Borrowers?

## Background & Premise {.smaller}

-   Bondora P2P Lending Service
-   European dataset spanning between 2009-2021 with 179,000+ individual loans
-   Data on 112 attributes of users individually and the characteristics of their loans

### Network Information

+-----------------+--------------------------------------------------------------------------------+
| **Type**        | **Undirected Bipartite**                                                       |
|                 |                                                                                |
| $\mathrm{B}_1$  | Borrowers identified by `UserName`                                             |
|                 |                                                                                |
| $\mathrm{B}_2$  | Reported `LoanUse`                                                             |
+-----------------+--------------------------------------------------------------------------------+
| **Edges**       | $\mathrm{B}_1 \rightarrow \mathrm{B}_2:$ User takes out specific kind of loan  |
|                 |                                                                                |
|                 | $\mathrm{B}_2 \rightarrow \mathrm{B}_1:$ Specific loan type belongs to a User  |
|                 |                                                                                |
|                 | $\mathrm{B}_1 \nleftrightarrow \mathrm{B}_2:$ Forbidden edges                  |
+-----------------+--------------------------------------------------------------------------------+

## Research Questions & Hypotheses {.smaller}

::: panel-tabset
### Study 1

-   **How do loan characteristics affect the frequency credit allocation among borrowers?**

-   Studied using **QAP Linear Regressions** since adjacency matrices are **valued**

+----------------------------------------------------------------------------------+-----------------------+-------------------------------------------------+
| Hypothesis                                                                       | Expected Result       | Specific Method                                 |
+==================================================================================+=======================+=================================================+
| H~1~^a^ : Users with the same loan use are likely to have the same credit rating | $\hat{\beta}_{1,1}>0$ | $Y=$ Adjacency matrix of shared loan use        |
|                                                                                  |                       |                                                 |
|                                                                                  |                       | $X_1=$ Adjacency matrix of shared credit rating |
|                                                                                  |                       |                                                 |
|                                                                                  |                       | $+$ $X_{2,3,..,n}=$ Controls                    |
+----------------------------------------------------------------------------------+-----------------------+-------------------------------------------------+

### Study 2

-   **How is the structure credit type allocation affected by borrower behaviour and its characteristics?**

-   Studied using **Bipartite ERGMs** with **binary edges**

+------------------------------------------------------------------------+---------------------+-----------------------------------------------------------------------------------------------------------------------+
| Hypothesis                                                             | Expected Result     | ERGM Term                                                                                                             |
+========================================================================+=====================+=======================================================================================================================+
| H~2~^a^ : Pairs of borrowers share same loan type                      | $\hat{\theta}>0$    | Endogenous + Dyad-dependent                                                                                           |
|                                                                        |                     |                                                                                                                       |
|                                                                        |                     | `gwb1dsp()`                                                                                                           |
+------------------------------------------------------------------------+---------------------+-----------------------------------------------------------------------------------------------------------------------+
| H~3~^a^ : Borrowers have a non-diverse loan use mix                    | $\hat{\theta}<0$    | Endogenous + Dyad-dependent                                                                                           |
|                                                                        |                     |                                                                                                                       |
|                                                                        |                     | `gwb1degree()`                                                                                                        |
+------------------------------------------------------------------------+---------------------+-----------------------------------------------------------------------------------------------------------------------+
| H~4~^a^ : Younger borrowers have a different loan use mix              | $\hat{\theta}\neq0$ | Exogenous + Dyad-independent                                                                                          |
|                                                                        |                     |                                                                                                                       |
|                                                                        |                     | `b1cov("b1_age")`                                                                                                     |
+------------------------------------------------------------------------+---------------------+-----------------------------------------------------------------------------------------------------------------------+
| H~5~^a^ : Borrowers of differing genders have different loan use mixes | $\hat{\theta}<0$    | Heterophily + Exogenous + Dyad-independent                                                                            |
|                                                                        |                     |                                                                                                                       |
|                                                                        |                     | `b1nodematch("b1_gender")`                                                                                            |
|                                                                        |                     |                                                                                                                       |
|                                                                        |                     | **Note:** measured indirectly through heterophily because `nodemix()` does not have an equivalent for bipartite ERGMs |
+------------------------------------------------------------------------+---------------------+-----------------------------------------------------------------------------------------------------------------------+
:::

## Literature {.smaller}

::: panel-tabset
### Study 1

**H~1~^a^ : Users with the same loan use are likely to have the same credit rating**

-   Studied by @serrano-cinca_determinants_2015

    -   Studies determinants of default with P2P lending data from LendingClub

    -   Identifies **Loan Purpose** as significant predictor of default

    -   Loans for weddings least risky but small business are riskiest

    -   $\Rightarrow$ Credit riskiness significantly associated with loan purpose

    -   $\Rightarrow$ If loan types vary in riskiness, then users should share creditworthiness

### Study 2

**H~2~^a^ : Pairs of borrowers share same loan type**

-   Studied by @liu_network_2024

    -   Studies impact of Bondora network centrality on credit risk

    -   Edges weighted by Gower distance based on borrower similarity characteristics

    -   Analyses degree centrality relative to defaulted vs non-defaulted loans

    -   Concludes that degree centrality significant factor in explaining credit risk

    -   $\Rightarrow$ If similarity between borrowers directly explains credit riskiness, then **users should cluster** around loans sharing similar risk profile [@serrano-cinca_determinants_2015]

**H~3~^a^ : Borrowers do not have a diverse loan use portfolio**

-   Studied by @lim_understanding_2023 and @abbasi_p2p_2021

    -   General literature review of P2P lending on variety of individuals, including businesses

    -   Across Malaysia, most common borrowing is for business financing by SMEs

    -   Millennials are the most populous user base of P2P lending platforms

    -   Authors argue business loans most popular because traditional commercial loans less accessible and more expensive in Malaysia

    -   *However*, Malaysian and SME context not directly applicable in Bondora's market

        -   @abbasi_p2p_2021 studies SME financing in OECD countries between 2011-2018 with P2P borrowing services

        -   Authors argue that OECD SMEs rely **less** on P2P because of greater accessibility of traditional loans

        -   $\Rightarrow$ Traditional European P2P users less likely to demand business loans $\rightarrow$ more diversity

**H~4~^a^ : Younger borrowers have a different loan use mix**

-   Explored by @cooper_consumption_2023

    -   Study of credit consumption patterns of younger individuals

    -   Authors find that younger individuals have *less access* to certain categories of traditional credit products and suggests that their credit portfolio structurally different than older borrowers

    -   Credit building fundamentally different in U.S., however, attitude towards credit type relevant

    -   $\Rightarrow$ If age is a factor in credit type, then we expect to see a **difference in loan use** across age

**H~5~^a^ : Borrowers of differing genders have different portfolios of loan use**

-   Explored by @aliano_role_2023 and @serrano-cinca_determinants_2015

    -   @aliano_role_2023 study of role of gender and education on probability of default with Bondora's dataset

    -   Authors find a consistent gender effect; women tend to have lower default rates on health, home, and business loans $\rightarrow$ structural gender differences in borrowing dynamics

        -   $\Rightarrow$ Given that loan purpose is directly related to loan default [@serrano-cinca_determinants_2015], we can expect that the **loan use mix has a gender effect**
:::

## Data Preprocessing {.smaller}

::: panel-tabset
### Procedures

-   Select subset of 112 attributes

``` r
c("LoanId", "UserName", "Age", "Gender","Country", "Amount", "Interest","LoanDuration","UseOfLoan", 
  "Education", "MaritalStatus","Rating", "Restructured", "MonthlyPayment")
```

-   Remove any loans with `NA` for any of the above features

-   Remove loans where `UseOfLoan == -1` to ensure complete data

-   Ensure dataset contains active users: keep `UserName` with 5+ loans

-   Create incidence matrix to construct partitions $\mathrm{B}_1$ and $\mathrm{B}_2$

-   **For ERGM models:**

    -   Create bipartite `network` object and add relevant vertex attributes

-   **For Linear QAP models:**

    -   Create adjacency matrices through matrix multiplication: $\mathbf{X}\cdot\mathbf{X}^{T}$

    -   Ensure no loops can exist by encoding all diagonals as zero

### Non-Network Descriptive Plots

```{r}
for (plt in list(h1, h2, h3, h4, h5, br1, br2)) {
  replayPlot(plt)
}
```
:::

## Basic Network Statistics {.smaller}

::: panel-tabset
### Network Visualisation

```{r}
replayPlot(ergm_plt)
```

### Descriptive Statistics Plot

```{r}
snafun::plot_centralities(net)
```
:::

## Study 1 \| Preliminary QAP Results {.smaller}

```{r include=FALSE}
var_names <- c("Intercept", "Rating","Loan Amount","Age","Gender",
               "Loan Duration","Restructured")
qap1_results <- qap_m1$coefficients
qap2_results <- qap_m2$coefficients
names(qap1_results) <- var_names
names(qap2_results) <- var_names
```

::: panel-tabset
### Notes

-   Linear QAP Regression used to study how borrowers with the same loan type choice are affected by having the same credit rating

-   The adjacency matrices used are **weighted** **by count** to give this study a level of detail we do not have with ERGM

    -   Because the weights vary significantly, **standardised** results are more meaningful to compare

-   Intercept has very significant impact: *not all effects can be accounted for and are missing*

### Parameters

+--------------------------+------------------------------------------------------------------+
| Variable Type            | Variable Name                                                    |
+==========================+==================================================================+
| **Dependent Variable**   | Weighted Adjacency Matrix of Borrowers with same Loan Type       |
+--------------------------+------------------------------------------------------------------+
| **Explanatory Variable** | Weighted Adjacency Matrix of Borrowers of the same Credit Rating |
+--------------------------+------------------------------------------------------------------+
| Control Variable         | Weighted Adjacency Matrix of Mean Loan Amount Differences        |
+--------------------------+------------------------------------------------------------------+
| Control Variable         | Weighted Adjacency Matrix of Age Differences                     |
+--------------------------+------------------------------------------------------------------+
| Control Variable         | Binary Adjacency Matrix of Homophily in Gender                   |
+--------------------------+------------------------------------------------------------------+
| Control Variable         | Weighted Adjacency Matrix of Mean Loan Duration Differences      |
+--------------------------+------------------------------------------------------------------+
| Control Variable         | Weighted Adjacency Matrix of Homophily in Loan Restructuring     |
+--------------------------+------------------------------------------------------------------+

### Specification & Results

``` r
qap_m1 <- sna::netlm(y = loan_use_mat,
                     x = list(rating_mat, amt_diffs_mat, age_diffs_mat,
                              gender_mat, loandur_diffs_mat, rest_mat),
                     nullhyp = "qapspp", reps = 2500)

# Standardised QAP Linear Regression
scaled_dep <- scale(loan_use_mat)
scaled_pred <- lapply(pred_vars, scale)

qap_m2 <- sna::netlm(y = scaled_dep,
                     x = scaled_pred,
                     nullhyp = "qapspp", reps = 2500)
```

```{r}
coef_table <- cbind(Unstd_Parameters = round(qap1_results, 4),
                    Ustd_t_vals = round(qap_m1$tstat, 2),
                    Std_Parameters = round(qap2_results, 4),
                    Std_t_vals = round(qap_m2$tstat, 2))
print(coef_table)
```

### Estimates Plot

```{r fig.width=20, fig.height=8}
replayPlot(qap_plt)
```
:::

## Study 2 \| ERGM \| `edges` {.smaller}

::: panel-tabset
### Summary

-   Sparse edges but bipartite, so take `edges` with grain of salt

-   Unconstrained, minimal model

-   GOF acceptable for an initial non-tuned ERGM

### Specification

``` r
formula_base_model <- bondora_net ~ edges
base_ergm <- ergm::ergm(formula_base_model)
```

```{r}
texreg::screenreg(list(base_ergm$model))
```

### GOF

```{r}
snafun::stat_plot_gof(base_ergm$gof)
```
:::

## Study 2 \| ERGM \| `gwb1degree` {.smaller}

::: panel-tabset
### Summary

**Modelling Notes**

-   GOF indicates that `b2degree` not modelled well

    -   Partition 2 degrees distributed very differently than Partition 1

-   Model 2 converges optimally with `MCMC.samplesize = 100000` and `MCMC.interval = 1000`

-   Greater model stability with `MCMC.steplength = 0.25`

-   No model constraints

**Results**

-   Highly positive and significant estimate for `gwb1degree` which is approximately `r odds(ergm_m1$model$coefficients[2])` in the odds scale

    -   This indicates that borrowers generally have a balanced loan use mix since it is almost equally likely the mix is either highly varied or very specific, relative to a network of random borrowers

-   Excellent MCMC diagnostics

### Specification

``` r
model_1_params <- bondora_net ~ edges + 
  # b1 decay can be very low since 9 b2
  gwb1degree(decay=0.15, fixed=TRUE) 

model_1 <- ergm::ergm(
  model_1_params,
  
  # Max b2 degree is 72, so this constraint is reasonable
  # and helps convergence significantly.
  # Technically in the Bondora population this can be 
  # far higher but we are studying a subsample.
  #constraints = ~ bd(minout = 0, maxout = 80),
  
  control = ergm::control.ergm(
    # Greater burn-in for cleaner result
    MCMC.burnin = 20000,
    # Greater sample size for greater stability
    MCMC.samplesize = 100000,
    seed = 42,
    MCMC.interval = 1000,
    # Only needed for convergence pvals to improve
    MCMLE.maxit = 45,
    # Smaller steps for stability
    MCMLE.steplength = 0.25,
    parallel = n_cores,
    parallel.type = "PSOCK"
  )
)
```

### Results

```{r}
texreg::screenreg(list(
  base_ergm$model, ergm_m1$model))
```

### MCMC

```{r fig.height=10, fig.width=10}
ergm::mcmc.diagnostics(ergm_m1$model)
```

### GOF

```{r}
snafun::stat_plot_gof(ergm_m1$gof)
```
:::

## Study 2 \| ERGM \| `gwb1dsp`  {.smaller}

::: panel-tabset
### Summary

**Modelling Notes**

-   Model constrained by minimum of zero `outdegrees` and maximum of 80

    -   Smaller simulation space reduces the noisiness of the MCMC mixing process significantly

    -   Maximum `b2degree=72` so 80 may be reasonable, but we understand the conceptual implications of this

-   `gwb1dsp` decay was set to 0.5 because there are far too many dyad-wise partners otherwise

    -   Cannot be set too high otherwise too many relevant dyad-wise partners are lost

**Results**

-   `gwb1dsp` estimate is highly significant and approximately `r odds(ergm_m2$model$coefficients[3])` in the odds scale

    -   This indicates that sets of two borrowers are less likely to share the same loan type compared to a random case where borrowers select their loan type randomly

-   MCMC diagnostics indicate good mixing, though `gwb1dsp` traces are right-skewed slightly

-   GOF acceptable, however, modelling `b2degree` is difficult

-   AIC/BIC highly negative as a result $\rightarrow$ Model 3 theoretically better than Model 2 given MCMC + GOF

    -   Extremely low Standard Error for `gwb1dsp`

### Specification

``` r
model_2_params <- bondora_net ~ edges + 
  # low decay important because there is high clustering around low degrees
  gwb1degree(decay=0.15, fixed=TRUE) + 
  # decay should be higher due to wider variation in degree but 
  # too high of degree makes the traces concentrated around the tails.
  gwb1dsp(decay=0.5, fixed=TRUE)      
  
model_2 <- ergm::ergm(
  model_2_params,
  
  # Max b2 degree is 72, so this constraint is reasonable
  # and helps convergence significantly.
  # Technically in the Bondora population this can be 
  # far higher but we are studying a subsample.
  constraints = ~ bd(minout = 0, maxout = 80),
  
  control = ergm::control.ergm(
    # Greater burn-in for cleaner result
    MCMC.burnin = 20000,
    # Greater sample size for greater stability
    MCMC.samplesize = 100000,
    seed = 42,
    MCMC.interval = 1000,
    # Only needed for convergence pvals to improve
    MCMLE.maxit = 45,
    # Smaller steps for stability
    MCMLE.steplength = 0.25,
    parallel = n_cores,
    parallel.type = "PSOCK"
    )
  )
```

### Results

```{r}
texreg::screenreg(list(base_ergm$model, ergm_m1$model, 
                       ergm_m2$model))
```

### MCMC

```{r fig.height=10, fig.width=10}
ergm::mcmc.diagnostics(ergm_m2$model)
```

### GOF

```{r}
snafun::stat_plot_gof(ergm_m2$gof)
```
:::

## Study 2 \| ERGM \| `b1nodematch("gender")` {.smaller}

::: panel-tabset
### Summary

**Modelling Notes**

-   Since `b1nodematch` is an exogenous dyad-independent term, the parameters related to convergence were mostly unaffected

**Results**

-   `b1nodematch` is used as a proxy for `nodemix` because it is unavailable for bipartite networks

-   `b1nodematch` significant at 5% and converted to the odds scale, estimate is `r odds(ergm_m3$model$coefficients[4])`

    -   This indicates that being of the same gender decreases the diversity of the borrower's requested loan types, compared to being of the other gender, all else equal

-   The inclusion of the `b1nodematch` term increased the strength of the `gwb1degree` term, but decreased `gwb1dsp`

-   The term decreased the significance of the `gwb1dsp` term, suggesting it may be less statistically relevant under this modelling context

-   MCMC diagnostics show that `b1nodematch` worsens convergence and skews the traces of the other terms, however, it is acceptable

-   GOF is similar to previous cases, however, the fit for `b1degree` is slightly less optimal

-   AIC suggests that this model offers less explanatory potential than the previous, however, BIC indicates the converse. Since the difference is small, this model is not seen as significantly better formulated

### Specification

``` r
model_3_params <- bondora_net ~ edges + 
  # low decay important because there is high clustering around low degrees
  gwb1degree(decay=0.15, fixed=TRUE) + 
  # decay should be higher due to wider variation in degree but 
  # too high of degree makes the traces concentrated around the tails.
  gwb1dsp(decay=0.5, fixed=TRUE) +
  # See differences across genders (implicitly, since b1nodemix unavailable)
  b1nodematch("b1_gender", diff=FALSE)

model_3 <- ergm::ergm(
  model_3_params,
  
  # Max b2 degree is 72, so this constraint is reasonable
  # and helps convergence significantly.
  # Technically in the Bondora population this can be 
  # far higher but we are studying a subsample.
  constraints = ~ bd(minout = 0, maxout = 80),
  
  control = ergm::control.ergm(
    # Greater burn-in for cleaner result
    MCMC.burnin = 20000,
    # Greater sample size for greater stability
    MCMC.samplesize = 100000,
    seed = 42,
    MCMC.interval = 1000,
    # Only needed for convergence pvals to improve
    MCMLE.maxit = 45,
    # Smaller steps for stability
    MCMLE.steplength = 0.25,
    parallel = n_cores,
    parallel.type = "PSOCK"
  )
)
```

### Results

```{r}
texreg::screenreg(list(base_ergm$model, ergm_m1$model, 
                       ergm_m2$model,ergm_m3$model))
```

### MCMC

```{r fig.height=10, fig.width=10}
ergm::mcmc.diagnostics(ergm_m3$model)
```

### GOF

```{r}
snafun::stat_plot_gof(ergm_m3$gof)
```
:::

## Study 2 \| ERGM \| `b1cov("age")` {.smaller}

::: panel-tabset
### Summary

**Modelling Notes**

-   Parameters related to convergence were unchanged since we have added an exogenous, dyad-independent term

**Results**

-   The `b1cov` term is statistically insignificant and the magnitude is zero, indicating that it is irrelevant to the explanatory power of this model

-   The inclusion of the `b1cov` term appears to stabilise the MCMC mixing process, improving the trace distributions

-   The GOF is not significantly different than previous iterations

-   BIC, which penalises model complexity more than AIC, is notably better than the previous models $\Rightarrow$ this model offers better statistical relevance than the previous models, given our subsample

### Specification

``` r
model_4_params <- bondora_net ~ edges + 
  # low decay important because there is high clustering around low degrees
  gwb1degree(decay=0.15, fixed=TRUE) + 
  # decay should be higher due to wider variation in degree but 
  # too high of degree makes the traces concentrated around the tails.
  gwb1dsp(decay=0.5, fixed=TRUE) +
  # See differences across genders (implicitly, since b1nodemix unavailable)
  b1nodematch("b1_gender", diff=TRUE) +
  # See if higher ages make a difference
  b1cov("b1_age")

model_4 <- ergm::ergm(
  model_4_params,
  
  # Max b2 degree is 72, so this constraint is reasonable
  # and helps convergence significantly.
  # Technically in the Bondora population this can be 
  # far higher but we are studying a subsample.
  constraints = ~ bd(minout = 0, maxout = 80),
  
  control = ergm::control.ergm(
    # Greater burn-in for cleaner result
    MCMC.burnin = 20000,
    # Greater sample size for greater stability
    MCMC.samplesize = 100000,
    seed = 42,
    MCMC.interval = 1000,
    # Only needed for convergence pvals to improve
    MCMLE.maxit = 45,
    # Smaller steps for stability
    MCMLE.steplength = 0.25,
    parallel = n_cores,
    parallel.type = "PSOCK"
  )
)
```

### Results

```{r}
texreg::screenreg(list(base_ergm$model, ergm_m1$model, 
                       ergm_m2$model, ergm_m3$model, 
                       ergm_m4$model))
```

### MCMC

```{r fig.height=10, fig.width=10}
ergm::mcmc.diagnostics(ergm_m4$model)
```

### GOF

```{r}
snafun::stat_plot_gof(ergm_m4$gof)
```
:::

## References

::: {#refs}
:::
