---
title: "Bondora Credit Allocation Study"
subtitle: "Progress Meeting 3"
author: "Group 7"
bibliography: references.bib
csl: apa-6th-edition.csl
institute: "Jheronimus Academy of Data Science"
execute:
  cache: true
format:
  revealjs: 
    margin: 0.05
    embed-resources: true
    self-contained-math: true 
    center-title-slide: true
    self-contained: true
    theme: simple 
    scrollable: true
    standalone: true
---

```{r include=FALSE}
#install.packages("here") # if you do not have this installed already

net <- readRDS(here::here("resources","objects","preprocessing","bondora_net.Rds"))

base_ergm <- readRDS(here::here("resources","objects","ergm","ergm_base_panel.Rds"))
ergm_m1 <- readRDS(here::here("resources","objects","ergm","ergm_m1_panel.Rds"))
ergm_m2 <- readRDS(here::here("resources","objects","ergm","ergm_m2_panel.Rds"))
ergm_m3 <- readRDS(here::here("resources","objects","ergm","ergm_m3_panel.Rds"))
ergm_m4 <- readRDS(here::here("resources","objects","ergm","ergm_m4_panel.Rds"))

qap_m1 <- readRDS(here::here("resources","objects","qap","qap_m1.Rds"))
qap_m2 <- readRDS(here::here("resources","objects","qap","qap_m2.Rds"))

h1 <- readRDS(here::here("resources","objects","preprocessing","hist_amt.Rds"))
h2 <- readRDS(here::here("resources","objects","preprocessing","hist_int.Rds"))
h3 <- readRDS(here::here("resources","objects","preprocessing","hist_loandur.Rds"))
h4 <- readRDS(here::here("resources","objects","preprocessing","hist_monpmt.Rds"))
h5 <- readRDS(here::here("resources","objects","preprocessing","hist_age.Rds"))

br1 <- readRDS(here::here("resources","objects","preprocessing","bar_loanuse.Rds"))
br2 <- readRDS(here::here("resources","objects","preprocessing","bar_rating.Rds"))

qap_plt <- readRDS(here::here("resources","objects","qap","unstd_std_plot.Rds"))
ergm_plt <- readRDS(here::here("resources","objects","ergm","network_plot.Rds"))
```

```{r include=FALSE}
prob <- function(input) {
  return(unname(round(exp(input)/(1+exp(input))),2))
}
odds <- function(input) {
  return(unname(round(exp(input)/(1+exp(input)),2)))
}
```

# How are Different Kinds of Credit Distributed among Bondora Borrowers?

## Background & Premise {.smaller}

-   Bondora P2P Lending Service
-   European dataset spanning between 2009-2021 with 179,000+ individual loans
-   Data on 112 attributes of users individually and the characteristics of their loans

### Network Information

+----------------+-------------------------------------------------------------------------------+
| **Type**       | **Undirected Bipartite**                                                      |
|                |                                                                               |
| $\mathrm{B}_1$ | Borrowers identified by `UserName`                                            |
|                |                                                                               |
| $\mathrm{B}_2$ | Reported `LoanUse`                                                            |
+----------------+-------------------------------------------------------------------------------+
| **Edges**      | $\mathrm{B}_1 \rightarrow \mathrm{B}_2:$ User takes out specific kind of loan |
|                |                                                                               |
|                | $\mathrm{B}_2 \rightarrow \mathrm{B}_1:$ Specific loan type belongs to a User |
|                |                                                                               |
|                | $\mathrm{B}_1 \nleftrightarrow \mathrm{B}_2:$ Forbidden edges                 |
+----------------+-------------------------------------------------------------------------------+

## Theoretical Premise and RQs {.smaller}

::: panel-tabset
### Overall Study

-   Traditional financial theory shows that investors exhibit behavioural biases in how they choose their investments [@aggarwal1981]

    -   Behaviour varies based on risk attitudes and information theory

-   Evidence shows that these same biases may also affect non-professional P2P borrowers [@lee_herding_2012; @herzenstein_strategic_2011]

-   Specifically, @ayal_behavioral_2018 show that there is insufficient attention on borrowers' portfolio formation

-   We want to understand how non professional P2P borrowers form their portfolios and how this is affected by behavioural biases

### Study 1

-   Investors' familiarity with certain assets can make investments seem less risky, driving them towards it [@ayal_behavioral_2018]

-   @galak_micro-finance_2010 show clustering among borrowers and lenders across demographic attributes

-   @amar_winning_2011 show that borrowers are averse to holding *large number* of debts

    ***Do borrowers vary in their loan use type and frequency based on their individual and loan characteristics?***

-   Linear QAP can assess similarities across borrowers, taking into account loan frequency

### LRQAP

@serrano-cinca_determinants_2015 show that heterogeneity in credit ratings explains significant differences across loan use purposes. Therefore:

**Hypothesis 1: Borrowers with the same credit ratings choose similar loan uses**

@ravina_love_2019 shows that occupation significant factor in whether loans are awarded. @ayal_behavioral_2018 suggests this can be generalised to reflect borrower preferences. Shared occupations can be indicative of similar preferences in loan use

**Hypothesis 2: Borrowers sharing the same occupation area tend to share loan uses**

### Study 2

-   Individuals' financial decisions are not independent [@kahneman_prospect_1979]

-   @ayal_behavioral_2018 shows that these non-independent preferences can influence other borrowers

-   Therefore, we investigate P2P behavioural biases from a structural perspective

***What behavioural mechanisms affect the structure of borrower-to-loan-use connections?***

### ERGM

@herzenstein_strategic_2011 shows that familiarity bias can exhibit itself through clustering, or "relational herding". Borrowers gravitate towards socially-validated or popular choices, resulting in overlapping decisions

**Hypothesis 3: Borrowers who share one loan purpose are likely to share another**

Theory of Perceived Diversification shows that borrowers' distress increases with perceived distinctiveness of each loan type [@ayal_perceived_2009]. Borrowers prefer to minimize this distress.

**Hypothesis 4: Borrowers exhibit preferences for minimal debt distinctiveness, selecting one loan use**

@cooper_consumption_2023 show that younger borrowers have fundamentally different credit use mix than older individuals due to how their access to credit shapes their consumption habits. Therefore:

**Hypothesis 5: Younger borrowers have different loan use mixes than older borrowers**

@aliano_role_2023 show a significant gender effect on different loan uses and @croson_gender_2009 explain this can be due to differences in risk perception and self-selection effects:

**Hypothesis 6: Gender differences lead to different loan use mixes**
:::

## Summary of Hypotheses {.smaller}

::: panel-tabset
### Study 1

-   LR QAP modelling effect of shared attributes on the frequency and type of same loan use shared

<!-- -->

-   $Y=$ Weighted adjacency matrix of shared loan use

+------------------------------------------------------------------------------+---------------------+----------------------------------------------------------+
| Hypothesis                                                                   | Expected Result     | Specific Method                                          |
+==============================================================================+=====================+==========================================================+
| H~1~^a^ : Borrowers with the same credit ratings choose similar loan uses    | $\hat{\beta}_{1}>0$ | $X_1=$ Weighted adjacency matrix of shared credit rating |
+------------------------------------------------------------------------------+---------------------+----------------------------------------------------------+
| H~2~^a^ : Borrowers sharing the same occupation area tend to share loan uses | $\hat{\beta}_{2}>0$ | $X_2=$ Binary adjacency matrix of shared occupation area |
+------------------------------------------------------------------------------+---------------------+----------------------------------------------------------+
| \-                                                                           | \-                  | $+$ $X_{2,3,..,n}=$ Controls                             |
+------------------------------------------------------------------------------+---------------------+----------------------------------------------------------+

### Study 2

-   Studied using **Bipartite ERGMs** with binary edges

+--------------------------------------------------------------------------------------------------+---------------------+---------------------------------+
| Hypothesis                                                                                       | Expected Result     | ERGM Term                       |
+==================================================================================================+=====================+=================================+
| H~3~^a^ : Borrowers who share one loan purpose are likely to share another                       | $\hat{\theta}>0$    | `cycle(4)`                      |
+--------------------------------------------------------------------------------------------------+---------------------+---------------------------------+
| H~4~^a^ : Borrowers exhibit preferences for minimal debt distinctiveness, selecting one loan use | $\hat{\theta}>0$    | `b1degree(1)`                   |
+--------------------------------------------------------------------------------------------------+---------------------+---------------------------------+
| H~5~^a^ : Younger borrowers have different loan use mixes than older borrowers                   | $\hat{\theta}\neq0$ | `b1cov("b1_age")`               |
+--------------------------------------------------------------------------------------------------+---------------------+---------------------------------+
| H~6~^a^ : Gender differences lead to different loan use mixes                                    | $\hat{\theta}\neq0$ | `b1factor("gender")`            |
+--------------------------------------------------------------------------------------------------+---------------------+---------------------------------+
:::

## Data Preprocessing {.smaller}

::: panel-tabset
### Procedures

-   Select subset of 112 attributes

``` r
keep_cols <- c("LoanId", "UserName","Age", "Gender", 
               "Country", "Amount", "Interest","LoanDuration",
              "UseOfLoan", "Rating", "Restructured", "MonthlyPayment",
              "OccupationArea", "BiddingStartedOn")
```

-   Remove any loans with `NA` for any of the above features

-   Most complete data is between 2014-2016, when `UseofLoan` was reported

    -   Filter `BiddingStartedOn` between these dates

-   Preserve original connections by not filtering any further. Instead, randomly sample 500 individuals to narrow dataset down

-   Create incidence matrix of $n \times m$, where $n$ is the unique borrowers and $m$ is the reported loan use to construct partitions $\mathrm{B}_1$ and $\mathrm{B}_2$

### Non-Network Descriptive Plots

```{r}
for (plt in list(h1, h2, h3, h4, h5, br1, br2)) {
  replayPlot(plt)
}
```
:::

## Basic Network Statistics {.smaller}

::: panel-tabset
### Network Visualisation

```{r}
replayPlot(ergm_plt)
```

### Descriptive Statistics Plot

```{r}
snafun::plot_centralities(net)
```
:::

## Study 1 \| Preliminary QAP Results {.smaller}

```{r include=FALSE}

```

::: panel-tabset
### Notes

-   Linear QAP Regression used to study how borrowers with the same loan type choice are affected by having the same credit rating

-   The adjacency matrices used are **weighted** **by count** to give this study a level of detail we do not have with ERGM

    -   Because the weights vary significantly, **standardised** results are more meaningful to compare

-   Intercept has very significant impact: *not all effects can be accounted for and are missing*

### Parameters

+--------------------------+------------------------------------------------------------------+
| Variable Type            | Variable Name                                                    |
+==========================+==================================================================+
| **Dependent Variable**   | Weighted Adjacency Matrix of Borrowers with same Loan Type       |
+--------------------------+------------------------------------------------------------------+
| **Explanatory Variable** | Weighted Adjacency Matrix of Borrowers of the same Credit Rating |
+--------------------------+------------------------------------------------------------------+
| Control Variable         | Weighted Adjacency Matrix of Mean Loan Amount Differences        |
+--------------------------+------------------------------------------------------------------+
| Control Variable         | Weighted Adjacency Matrix of Age Differences                     |
+--------------------------+------------------------------------------------------------------+
| Control Variable         | Binary Adjacency Matrix of Homophily in Gender                   |
+--------------------------+------------------------------------------------------------------+
| Control Variable         | Weighted Adjacency Matrix of Mean Loan Duration Differences      |
+--------------------------+------------------------------------------------------------------+
| Control Variable         | Weighted Adjacency Matrix of Homophily in Loan Restructuring     |
+--------------------------+------------------------------------------------------------------+

### Specification & Results

``` r
qap_m1 <- sna::netlm(y = loan_use_mat,
                     x = list(rating_mat, amt_diffs_mat, age_diffs_mat,
                              gender_mat, loandur_diffs_mat, rest_mat),
                     nullhyp = "qapspp", reps = 2500)

# Standardised QAP Linear Regression
scaled_dep <- scale(loan_use_mat)
scaled_pred <- lapply(pred_vars, scale)

qap_m2 <- sna::netlm(y = scaled_dep,
                     x = scaled_pred,
                     nullhyp = "qapspp", reps = 2500)
```

```{r}

```

### Estimates Plot

```{r fig.width=20, fig.height=8}
replayPlot(qap_plt)
```
:::

## Study 2 \| ERGM \| `edges` {.smaller}

::: panel-tabset
### Summary

### Specification

``` r
formula_base_model <- bondora_net ~ edges
base_ergm <- ergm::ergm(formula_base_model)
```

```{r}
texreg::screenreg(list(base_ergm$model))
```

### GOF

```{r}
snafun::stat_plot_gof(base_ergm$gof)
```
:::

## Study 2 \| ERGM \| `b1degree(1)` {.smaller}

::: panel-tabset
### Summary

**Modelling Notes**

**Results**

### Specification

``` r
```

### Results

```{r}
texreg::screenreg(list(
  base_ergm$model, ergm_m1$model))
```

### MCMC

```{r fig.height=10, fig.width=10}
ergm::mcmc.diagnostics(ergm_m1$model)
```

### GOF

```{r}
snafun::stat_plot_gof(ergm_m1$gof)
```
:::

## Study 2 \| ERGM \| `cycle(4)` {.smaller}

::: panel-tabset
### Summary

**Modelling Notes**

**Results**

### Specification

``` r
```

### Results

```{r}
texreg::screenreg(list(base_ergm$model, ergm_m1$model, 
                       ergm_m2$model))
```

### MCMC

```{r fig.height=10, fig.width=10}
ergm::mcmc.diagnostics(ergm_m2$model)
```

### GOF

```{r}
snafun::stat_plot_gof(ergm_m2$gof)
```
:::

## Study 2 \| ERGM \| `b1factor("gender")` {.smaller}

::: panel-tabset
### Summary

**Modelling Notes**

**Results**

### Specification

``` r
```

### Results

```{r}
texreg::screenreg(list(base_ergm$model, ergm_m1$model, 
                       ergm_m2$model,ergm_m3$model))
```

### MCMC

```{r fig.height=10, fig.width=10}
ergm::mcmc.diagnostics(ergm_m3$model)
```

### GOF

```{r}
snafun::stat_plot_gof(ergm_m3$gof)
```
:::

## Study 2 \| ERGM \| `b1cov("age")` {.smaller}

::: panel-tabset
### Summary

**Modelling Notes**

**Results**

### Specification

``` r
```

### Results

```{r}
texreg::screenreg(list(base_ergm$model, ergm_m1$model, 
                       ergm_m2$model, ergm_m3$model, 
                       ergm_m4$model))
```

### MCMC

```{r fig.height=10, fig.width=10}
ergm::mcmc.diagnostics(ergm_m4$model)
```

### GOF

```{r}
snafun::stat_plot_gof(ergm_m4$gof)
```
:::

## References

::: {#refs}
:::
