---
title: "Bondora Credit Allocation Study"
subtitle: "Progress Meeting 3"
author: "Group 7"
bibliography: references.bib
csl: apa-6th-edition.csl
institute: "Jheronimus Academy of Data Science"
execute:
  cache: true
format:
  revealjs: 
    margin: 0.05
    embed-resources: true
    self-contained-math: true 
    center-title-slide: true
    self-contained: true
    theme: simple 
    scrollable: true
    standalone: true
---

```{r include=FALSE}
#install.packages("here") # if you do not have this installed already

net <- readRDS(here::here("resources","objects","preprocessing","bondora_net.Rds"))
```

```{r include=FALSE}
prob <- function(input) {
  return(unname(round(exp(input)/(1+exp(input))),3))
}
odds <- function(input) {
  return(unname(round(exp(input)/(1+exp(input)),3)))
}
```

# How are Different Kinds of Credit Distributed among Bondora Borrowers?

## Background & Premise {.smaller}

-   Bondora P2P Lending Service
-   European dataset spanning between 2009-2021 with 179,000+ individual loans
-   Data on 112 attributes of users individually and the characteristics of their loans

### Network Information

+----------------+-------------------------------------------------------------------------------+
| **Type**       | **Undirected Bipartite**                                                      |
|                |                                                                               |
| $\mathrm{B}_1$ | Borrowers identified by `UserName`                                            |
|                |                                                                               |
| $\mathrm{B}_2$ | Reported `LoanUse`                                                            |
+----------------+-------------------------------------------------------------------------------+
| **Edges**      | $\mathrm{B}_1 \rightarrow \mathrm{B}_2:$ User takes out specific kind of loan |
|                |                                                                               |
|                | $\mathrm{B}_2 \rightarrow \mathrm{B}_1:$ Specific loan type belongs to a User |
|                |                                                                               |
|                | $\mathrm{B}_1 \nleftrightarrow \mathrm{B}_2:$ Forbidden edges                 |
+----------------+-------------------------------------------------------------------------------+

## Theoretical Premise and RQs {.smaller}

::: panel-tabset
### Overall Study

-   Traditional financial theory shows that investors exhibit behavioural biases in how they choose their investments [@aggarwal1981]

    -   Behaviour varies based on risk attitudes and information theory

-   Evidence shows that these same biases may also affect non-professional P2P borrowers [@lee_herding_2012; @herzenstein_strategic_2011]

-   Specifically, @ayal_behavioral_2018 show that there is insufficient attention on borrowers' portfolio formation

-   We want to understand how non professional P2P borrowers form their portfolios and how this is affected by behavioural biases

### Study 1

-   Investors' familiarity with certain assets can make investments seem less risky, driving them towards it [@ayal_behavioral_2018]

-   @galak_micro-finance_2010 show clustering among borrowers and lenders across demographic attributes

-   @amar_winning_2011 show that borrowers are averse to holding *large number* of debts

    ***Do borrowers vary in their loan use type and frequency based on their individual and loan characteristics?***

-   Linear QAP can assess similarities across borrowers, taking into account loan frequency

### LRQAP

@serrano-cinca_determinants_2015 show that heterogeneity in credit ratings explains significant differences across loan use purposes. Therefore:

**Hypothesis 1: Borrowers with the same credit ratings choose similar loan uses**

@ravina_love_2019 shows that occupation significant factor in whether loans are awarded. @ayal_behavioral_2018 suggests this can be generalised to reflect borrower preferences. Shared occupations can be indicative of similar preferences in loan use

**Hypothesis 2: Borrowers sharing the same occupation area tend to share loan uses**

### Study 2

-   Individuals' financial decisions are not independent [@kahneman_prospect_1979]

-   @ayal_behavioral_2018 shows that these non-independent preferences can influence other borrowers

-   Therefore, we investigate P2P behavioural biases from a structural perspective

***What behavioural mechanisms affect the structure of borrower-to-loan-use connections?***

### ERGM

@herzenstein_strategic_2011 shows that familiarity bias can exhibit itself through clustering, or "relational herding". Borrowers gravitate towards socially-validated or popular choices, resulting in overlapping decisions

**Hypothesis 3: Borrowers who share one loan purpose are likely to share another**

Theory of Perceived Diversification shows that borrowers' distress increases with perceived distinctiveness of each loan type [@ayal_perceived_2009]. Borrowers prefer to minimize this distress.

**Hypothesis 4: Borrowers exhibit preferences for minimal debt distinctiveness, selecting one loan use**

@cooper_consumption_2023 show that younger borrowers have fundamentally different credit use mix than older individuals due to how their access to credit shapes their consumption habits. Therefore:

**Hypothesis 5: Younger borrowers have different loan use mixes than older borrowers**

@aliano_role_2023 show a significant gender effect on different loan uses and @croson_gender_2009 explain this can be due to differences in risk perception and self-selection effects:

**Hypothesis 6: Gender differences lead to different loan use mixes**
:::

## Summary of Hypotheses {.smaller}

::: panel-tabset
### Study 1

-   LR QAP modelling effect of shared attributes on the frequency and type of same loan use shared

<!-- -->

-   $Y=$ Weighted adjacency matrix of shared loan use

+------------------------------------------------------------------------------+---------------------+----------------------------------------------------------+
| Hypothesis                                                                   | Expected Result     | Specific Method                                          |
+==============================================================================+=====================+==========================================================+
| H~1~^a^ : Borrowers with the same credit ratings choose similar loan uses    | $\hat{\beta}_{1}>0$ | $X_1=$ Weighted adjacency matrix of shared credit rating |
+------------------------------------------------------------------------------+---------------------+----------------------------------------------------------+
| H~2~^a^ : Borrowers sharing the same occupation area tend to share loan uses | $\hat{\beta}_{2}>0$ | $X_2=$ Binary adjacency matrix of shared occupation area |
+------------------------------------------------------------------------------+---------------------+----------------------------------------------------------+
| \-                                                                           | \-                  | $+$ $X_{2,3,..,n}=$ Controls                             |
+------------------------------------------------------------------------------+---------------------+----------------------------------------------------------+

### Study 2

-   Studied using **Bipartite ERGMs** with binary edges

+--------------------------------------------------------------------------------------------------+---------------------+---------------------------------+
| Hypothesis                                                                                       | Expected Result     | ERGM Term                       |
+==================================================================================================+=====================+=================================+
| H~3~^a^ : Borrowers who share one loan purpose are likely to share another                       | $\hat{\theta}>0$    | `cycle(4)`                      |
+--------------------------------------------------------------------------------------------------+---------------------+---------------------------------+
| H~4~^a^ : Borrowers exhibit preferences for minimal debt distinctiveness, selecting one loan use | $\hat{\theta}>0$    | `b1degree(1)`                   |
+--------------------------------------------------------------------------------------------------+---------------------+---------------------------------+
| H~5~^a^ : Younger borrowers have different loan use mixes than older borrowers                   | $\hat{\theta}\neq0$ | `b1cov("b1_age")`               |
+--------------------------------------------------------------------------------------------------+---------------------+---------------------------------+
| H~6~^a^ : Gender differences lead to different loan use mixes                                    | $\hat{\theta}\neq0$ | `b1factor("gender")`            |
+--------------------------------------------------------------------------------------------------+---------------------+---------------------------------+
:::

## Data Preprocessing {.smaller}

::: panel-tabset
### Procedures

-   Select subset of 112 attributes

``` r
keep_cols <- c("LoanId", "UserName","Age", "Gender", 
               "Country", "Amount", "Interest","LoanDuration",
              "UseOfLoan", "Rating", "Restructured", "MonthlyPayment",
              "OccupationArea", "BiddingStartedOn")
```

-   Remove any loans with `NA` for any of the above features

-   Most complete data is between 2014-2016, when `UseofLoan` was reported (see user counts plot)

    -   Filter `BiddingStartedOn` between these dates

-   Preserve original connections by not filtering any further. Instead, randomly sample 500 individuals to narrow dataset down

-   Create incidence matrix of $n \times m$, where $n$ is the unique borrowers and $m$ is the reported loan use to construct partitions $\mathrm{B}_1$ and $\mathrm{B}_2$

### User Counts Plot

```{r fig.width=20, fig.height=8}
replayPlot(readRDS(here::here("resources","objects","preprocessing",
                              "user_cnt_plt.Rds")))
```

### Non-Network Descriptive Plots

```{r fig.width=20, fig.height=8}
#| warning: false

h1 <- readRDS(here::here("resources","objects","preprocessing",
                         "hist_amt.Rds"))
h2 <- readRDS(here::here("resources","objects","preprocessing",
                         "hist_int.Rds"))
h3 <- readRDS(here::here("resources","objects","preprocessing",
                         "hist_loandur.Rds"))
h4 <- readRDS(here::here("resources","objects","preprocessing",
                         "hist_monpmt.Rds"))
h5 <- readRDS(here::here("resources","objects","preprocessing",
                         "hist_age.Rds"))

br1 <- readRDS(here::here("resources","objects","preprocessing",
                          "bar_loanuse.Rds"))
br2 <- readRDS(here::here("resources","objects","preprocessing",
                          "bar_rating.Rds"))
br3 <- readRDS(here::here("resources","objects","preprocessing",
                          "bar_occupation.Rds"))

for (plt in list(h1, h2, h3, h4, h5, br1, br2, br3)) {
  replayPlot(plt)
}
```
:::

## Basic Network Statistics {.smaller}

::: panel-tabset
### Network Visualisation

```{r fig.width=20, fig.height=8}
#| warning: false
net_plt <- readRDS(here::here("resources","objects","ergm","network_plot.Rds"))

replayPlot(net_plt)
```

### Descriptive Statistics

```{r}
knitr::kable(readRDS(here::here("resources","objects","ergm",
                                "net_summary.Rds")))
```

### Descriptive Statistics Plot

```{r fig.width=20, fig.height=8}
snafun::plot_centralities(net)
```
:::

## Study 1 \| QAP Results {.smaller}

::: panel-tabset
### Notes

-   Linear QAP Regression used to study how borrowers with the same loan type choice are affected by having the same credit rating

-   The adjacency matrices used are **weighted** **by count** to give this study a level of detail we do not have with ERGM

    -   Because the weights vary significantly, **standardised** results are more meaningful to compare

-   Intercept has very significant impact: *not all effects can be accounted for and are missing*

### Specification

``` r
# Make function to run QAPs automatically
run_QAP <- function(y_mat, xlist, varnames, model_name) {
  
  # Run QAP Linear Regression
  model <- sna::netlm(y = y_mat, x = xlist, nullhyp = "qapspp", reps = 1000)
  model$names <- varnames
  summary(model)
  
  # Obtain results from Model and Name them
  results <- model$coefficients
  names(results) <- varnames
  
  # Add significance stars to model results
  results_sig <- paste(round(results,3), t_to_stars(model$tstat))
  
  # Save Model
  saveRDS(model, here::here("resources","objects","qap",
                            paste0(model_name,".Rds")))
  
  # Prepare function output
  list_names <- c("model","results","results_sig")
  items <- list(model, results, results_sig)
  names(items) <- list_names
  
  return(items)
}

# -----------------------------------------------------------------------------#

# Basic QAP Linear Regression 1 - Unstandardised + No Controls
qap_m1 <- run_QAP(loan_use_mat, main_pred_vars, main_pred_names, "qap_m1")

# Basic QAP Linear Regression 1 - Standardised + No Controls
scaled_dep <- scale(loan_use_mat)
scaled_pred <- lapply(main_pred_vars, scale)

qap_m2 <- run_QAP(scaled_dep, scaled_pred, main_pred_names, "qap_m2")

# Basic QAP Linear Regression 2 - Unstandardised + Controls
qap_m3 <- run_QAP(loan_use_mat, all_pred_vars, var_names, "qap_m3")

# Basic QAP Linear Regression 2 - Standardised + Controls
scaled_pred_2 <- lapply(all_pred_vars, scale)

qap_m4 <- run_QAP(scaled_dep, scaled_pred_2, var_names, "qap_m4")
```

### Results

```{r include=FALSE}
qap_m1 <- readRDS(here::here("resources","objects","qap","qap_m1.Rds"))
qap_m2 <- readRDS(here::here("resources","objects","qap","qap_m2.Rds"))
qap_m3 <- readRDS(here::here("resources","objects","qap","qap_m3.Rds"))
qap_m4 <- readRDS(here::here("resources","objects","qap","qap_m4.Rds"))

diag_1 <- readRDS(here::here("resources","objects","qap","qap_m1_diags.Rds"))
diag_2 <- readRDS(here::here("resources","objects","qap","qap_m2_diags.Rds"))
diag_3 <- readRDS(here::here("resources","objects","qap","qap_m3_diags.Rds"))
diag_4 <- readRDS(here::here("resources","objects","qap","qap_m4_diags.Rds"))

qap_table_items <- readRDS(here::here("resources","objects","qap",                                 "qap_tables.Rds"))
```

```{r}
#| asis: true
texreg::screenreg(qap_table_items[[1]], 
               custom.model.names = qap_table_items[[2]],
               digits = 3)
```

### Estimates Plot

```{r fig.width=20, fig.height=8}
#| warning: false
qap_nocntrl_plt <- readRDS(here::here("resources","objects","qap",
                                      "qap_nocontrols_plots.Rds"))
qap_cntrl_plt <- readRDS(here::here("resources","objects","qap",
                              "qap_controls_plots.Rds"))
replayPlot(qap_nocntrl_plt)
replayPlot(qap_cntrl_plt)
```

### Diagnostics

```{r fig.width=20, fig.height=8}
replayPlot(diag_1)
replayPlot(diag_2)
replayPlot(diag_3)
replayPlot(diag_4)
```
:::

## Study 2 \| ERGM \| `edges` {.smaller}

::: panel-tabset
### Summary

```{r include=FALSE}
base_ergm <- readRDS(here::here("resources","objects","ergm",
                                "ergm_base_panel.Rds"))
```

-   Significant negative coefficient indicates large amount of sparsity

-   Low SE

-   Loan use mix is likely not random

-   $\Rightarrow$ Base ERGM model looks as expected

### Specification

``` r
# Base Model + GOF
formula_base_model <- bondora_net ~ edges
base_ergm <- ergm::ergm(formula_base_model)
```

```{r}
texreg::screenreg(list(base_ergm$model))
```

### GOF

```{r}
snafun::stat_plot_gof(base_ergm$gof)
```
:::

## Study 2 \| ERGM \| `b1degree(1)` {.smaller}

::: panel-tabset
### Summary

```{r include=FALSE}
ergm_m1 <- readRDS(here::here("resources","objects","ergm","ergm_m1_panel.Rds"))
```

**Modelling Notes**

-   Quick convergence within 5-10 iterations

**Results**

-   Good MCMC + GOF

-   Significant and positive coefficient `r round(ergm_m1$model$coefficients[2],3)` translates into probability of `r prob(ergm_m1$model$coefficients[2])`.

-   Low SE

-   Increases `edges` coefficient

-   Significantly better AIC & BIC in Model 2

    -   Given MCMC + GOF, considered better model

-   $\Rightarrow$ Individuals generally choose single loan use categories and modelling this improves the model significantly over the base ERGM

### Specification

``` r
# Iteration 1 + MCMC Diagnostics + GOF
model_1_params <- bondora_net ~ edges + 
  # See if individuals tend to pick one loan type
  b1degree(1) 

model_1 <- ergm::ergm(
  model_1_params,
  #constraints = ~ bd(minout = 0, maxout = 80),
  
  control = ergm::control.ergm(
    # Greater burn-in for cleaner result
    MCMC.burnin = 20000,
    # Greater sample size for greater stability
    MCMC.samplesize = 100000,
    seed = 42,
    MCMC.interval = 1000,
    # Only needed for convergence pvals to improve
    MCMLE.maxit = 45,
    # Smaller steps for stability
    MCMLE.steplength = 0.25,
    parallel = n_cores,
    parallel.type = "PSOCK"
  )
)
```

### Results

```{r}
texreg::screenreg(list(base_ergm$model, ergm_m1$model))
```

### MCMC

```{r fig.height=10, fig.width=10}
ergm::mcmc.diagnostics(ergm_m1$model)
```

### GOF

```{r}
snafun::stat_plot_gof(ergm_m1$gof)
```
:::

## Study 2 \| ERGM \| `cycle(4)` {.smaller}

::: panel-tabset
### Summary

```{r include=FALSE}
ergm_m2 <- readRDS(here::here("resources","objects","ergm","ergm_m2_panel.Rds"))
```

**Modelling Notes**

-   Quick convergence within 5-10 iterations

**Results**

-   Good MCMC + GOF

-   Insignificant estimate, but low SE

-   Lowers `edges` quite a lot

-   Does not impact `b1degree(1)` estimate magnitude large amount

-   No significant change in AIC/BIC between Model 2 to Model 3

-   $\Rightarrow$ Individuals have common loan uses among them and the term's inclusion improves the model significantly

### Specification

``` r
# Iteration 2 + MCMC Diagnostics + GOF
model_2_params <- bondora_net ~ edges + b1degree(1) + 
  # See if there is clustering around pairs of loan types
  cycle(4)
  
model_2 <- ergm::ergm(
  model_2_params,
  #constraints = ~ bd(minout = 0, maxout = 80),
  
  control = ergm::control.ergm(
    # Greater burn-in for cleaner result
    MCMC.burnin = 20000,
    # Greater sample size for greater stability
    MCMC.samplesize = 100000,
    seed = 42,
    MCMC.interval = 1000,
    # Only needed for convergence pvals to improve
    MCMLE.maxit = 45,
    # Smaller steps for stability
    MCMLE.steplength = 0.25,
    parallel = n_cores,
    parallel.type = "PSOCK"
    )
  )
```

### Results

```{r}
texreg::screenreg(list(base_ergm$model, ergm_m1$model, 
                       ergm_m2$model))
```

### MCMC

```{r fig.height=10, fig.width=10}
ergm::mcmc.diagnostics(ergm_m2$model)
```

### GOF

```{r}
snafun::stat_plot_gof(ergm_m2$gof)
```
:::

## Study 2 \| ERGM \| `b1factor("gender")` {.smaller}

::: panel-tabset
### Summary

```{r include=FALSE}
ergm_m3 <- readRDS(here::here("resources","objects","ergm","ergm_m3_panel.Rds"))
```

**Modelling Notes**

-   Originally, we wanted to test `b1nodematch("gender")`, however this fails to mix

-   Gender effect can be indirectly observed with `b1factor("gender")` as well

    -   Since this converges without issue, we chose this as a replacement term

-   Term change does not impact original hypothesis significantly

**Results**

-   Good MCMC + GOF

-   Gender term insignificant and with low SE

-   AIC/BIC not changed significantly

-   $\Rightarrow$ Gender does not impact loan use choice significantly and the model does not significantly improve explanatory power

### Specification

``` r
# Iteration 3 + MCMC Diagnostics + GOF
model_3_params <- bondora_net ~ edges + b1degree(1) + cycle(4) +
  
  # See if there is gender effect
  b1factor("b1_gender")

model_3 <- ergm::ergm(
  model_3_params,
  #constraints = ~ bd(minout = 0, maxout = 80),
  
  control = ergm::control.ergm(
    # Greater burn-in for cleaner result
    # Cycles mixed with the rest of the terms requires more burnin samples
    MCMC.burnin = 20000,
    # Greater sample size for greater stability
    MCMC.samplesize = 100000,
    seed = 42,
    # Reducing interval to reduce complexity between intervals
    MCMC.interval = 1000,
    # Only needed for convergence pvals to improve
    MCMLE.maxit = 25,
    # Smaller steps for stability
    MCMLE.steplength = 0.25,
    parallel = n_cores,
    parallel.type = "PSOCK"
  )
)
```

### Results

```{r}
texreg::screenreg(list(base_ergm$model, ergm_m1$model, 
                       ergm_m2$model,ergm_m3$model))
```

### MCMC

```{r fig.height=10, fig.width=10}
ergm::mcmc.diagnostics(ergm_m3$model)
```

### GOF

```{r}
snafun::stat_plot_gof(ergm_m3$gof)
```
:::

## Study 2 \| ERGM \| `b1cov("age")` {.smaller}

::: panel-tabset
### Summary

```{r include=FALSE}
ergm_m4 <- readRDS(here::here("resources","objects","ergm","ergm_m4_panel.Rds"))
```

**Modelling Notes**

-   No issues with convergence

**Results**

-   Good MCMC + GOF

-   Insignificant term but with low SE

-   Increases `edges` term more strongly than model with just `b1factor("gender")`

-   $\Rightarrow$ Age is not a significant factor in the formation of loan use mix and its inclusion does not improve model's explanatory power significantly

### Specification

``` r
# Iteration 4 + MCMC Diagnostics + GOF
model_4_params <- bondora_net ~ edges + b1degree(1) + cycle(4) +
  b1factor("b1_gender") +
  
  # See if higher ages make a difference
  b1cov("b1_age")

model_4 <- ergm::ergm(
  model_4_params,
  #constraints = ~ bd(minout = 0, maxout = 80),
  
  control = ergm::control.ergm(
    # Greater burn-in for cleaner result
    MCMC.burnin = 20000,
    # Greater sample size for greater stability
    MCMC.samplesize = 100000,
    seed = 42,
    MCMC.interval = 1000,
    # Only needed for convergence pvals to improve
    MCMLE.maxit = 45,
    # Smaller steps for stability
    MCMLE.steplength = 0.25,
    parallel = n_cores,
    parallel.type = "PSOCK"
  )
)
```

### Results

```{r}
texreg::screenreg(list(base_ergm$model, ergm_m1$model, 
                       ergm_m2$model, ergm_m3$model, 
                       ergm_m4$model))
```

### MCMC

```{r fig.height=10, fig.width=10}
ergm::mcmc.diagnostics(ergm_m4$model)
```

### GOF

```{r}
snafun::stat_plot_gof(ergm_m4$gof)
```
:::

## References

::: {#refs}
:::
